---
output: html_document
params:
  sims:
    label: "Simulations"
    value: 1000
    input: numeric
    min: 1000
    max: 1000000
    step: 1000
  abun:
    label: "Total Abundance"
    value: 48000
    input: slider
    min: 1000
    max: 500000
    step: 100
  fracFem:
    label: "Fraction Females"
    value: 0.50
    input: slider
    min: 0.20
    max: 0.80
    step: 0.05
  mature:
    label: "Maturity"
    value: 0.15
    input: numeric
    min: 0.10
    max: 0.25
    step: 0.05
  recruit:
    label: "Recruitment"
    value: 5
    input: slider
    min: 1
    max: 20
    step: 1
  agesHarvest:
    label: "Ages (harvest)"
    value: !r c(10, 15)
    input: slider
    min: 5
    max: 20
    step: 1
  surv:
    label: "Survival Rate"
    value: !r NA
    input: numeric
    min: 0.10
    max: 0.95
    step: 0.05
  harv:
    label: "Harvest Rate"
    value: !r NA
    input: numeric
    min: 0
    max: 0.30
    step: 0.01
  age0:
    label: "Age-0"
    value: !r NA
    input: numeric
    min: 60000000
    max: 250000000
    step: 1000000
  d:
    label: "Age Data"
    value: trammel_catch
    input: select
    choices: [trammel_catch]
---

```{r setup, include=FALSE}

knitr::opts_knit$set(root.dir = "~/RProjects/SturgeonPopMetrics/")

```

## Introduction


## Libraries

We load the `spopmodel` package, currently available on GitHub. For now (`r Sys.Date()`), this is the only package required.

```{r load-libraries} 

library(spopmodel)
# library(package)

```

```{r load-data}

# looking for an easier way to use internal data from `params`; for now this
# will suffice might needt to implement `if` or `switch` to select other
# datasets (05-Aug-2019)

# data(params$d, package = "spopmodel")
d <- get(
  params[["d"]],
  pos = which(search() %in% "package:spopmodel")
)

```

```{r fish-gte190, eval=TRUE}

bool_gte190 <- d[["FL"]] >= 190

bool_ageNA <- is.na(d[["Age"]])

sum(bool_gte190 & bool_ageNA)

d[bool_gte190 & bool_ageNA, "Age"] <- 19

# verification
# d[bool_gte190, ]

# clean up
rm(bool_gte190, bool_ageNA) 

```

#### Length Frequency

```{r len-freq}

len_freq <- Frequency(d[["FL"]], binWidth = 5)

unlist(len_freq$xstats())

```

```{r plot-len_freq}

plot(len_freq, xlab = "Length")

```

#### Age Frequency

```{r age-freq}

ages <- AgeEach(
  data = d,
  len  = FL,
  age  = Age,
  lenBreaks = len_freq[["breaks"]]
)

age_freq <- table(ages[["Ages"]], dnn = NULL)

# as a possible check
# sum(age_freq)
age_freq

```

#### Mean Length-at-Age

```{r mean-len-age}

# get mean-length-at-age given lengths with assigned ages
mean_len_age <- aggregate(
  d[["FL"]],
  # by = list(Age = d$Age),
  by = d["Age"],
  FUN = mean
)

# for convenience
colnames(mean_len_age)[2] <- "MeanFL"

# uncomment to display if desired
# mean_len_age

```

```{r plot-mla}

# to diplay length given age along with mean-length-at-age

plot(
  x = jitter(d[["Age"]]),
  y = d[["FL"]],
  pch = 19,
  col = "grey70",
  las = 1,
  ylab = "Length",
  xlab = "Age"
)

points(mean_len_age, pch = "+", col = "darkred")

```

#### Age Distribution

```{r age-dist}

# establishes age distribution given input from params

age_distribution <- AgeDist(
  ageFreq = age_freq,
  abund = params[["abun"]],
  fracFemale = params[["fracFem"]]
)

# for later analyses
est_abundance <- age_distribution[["EstAgeAbund"]]

```

```{r log-lin-abun}

yval <- log(as.vector(est_abundance))
xval <- as.numeric(names(est_abundance))

# log-linear abundance to predict age 1-2 abundance
mod <- lm(yval ~ xval)

```

```{r plot-log-lin-abun}

plot(
  x = xval,
  y = yval,
  type = "b", 
  col='grey60',
  xlab = "Age",
  ylab = "log(EstAbund)",
  panel.first = abline(mod, col = "blue")
)

# clean up
rm(xval, yval)

```

Use model to predict younger ages

```{r predict-young-fish}

age1_2 <- exp(predict(object = mod, newdata = list(xval = c(1, 2))))

females <- c(
  age1_2 * age_distribution[["FracFemale"]],
  age_distribution[["CountFemByAge"]]
)

```

#### Spawning Probability

```{r mean-age-vars}

# subset of age > 9
bool_mlaa_gt9 <- mean_len_age[["Age"]] > 9  

mean_len <- mean_len_age[bool_mlaa_gt9, "MeanFL"]
age_at_len <- mean_len_age[bool_mlaa_gt9, "Age"]

```

```{r p-spawn}

p_spawn <- SpawningProb(
  len = mean_len,
  age = age_at_len,
  mature = params[["mature"]]
)

# create dataframe to be used in simulations
prob_spawn2 <- rbind(
  data.frame(Age = 0:9, Prob = 0, Err = 0),
  p_spawn[, c("Age", "Prob", "Err")]
)

```

#### Age Distribution (complete)

```{r age-0}

# Devore's (et al. 1995) equation to calculate number of eggs based on fork
# length

eggs_female <- 0.072 * mean_len^2.94

age0 <- sum(p_spawn[["Prob"]] * females[10:19] * eggs_female)

if (!is.na(params[["age0"]])) age0 <- params[["age0"]]

```

```{r age-dist-complete}

age_dist2 <- data.frame(
  Age = as.numeric(c(0, names(females))),
  Freq = c(age0, unname(females)),
  row.names = NULL
)

```

#### Egg Count

```{r num-eggs}

# egg count age-10 to age-19
egg_count <- EggCount(len = mean_len, age = age_at_len)

num_eggs <- rbind(
  data.frame(Age = 0:9, Count = 0, Err = 0),
  egg_count[, c("Age", "Count", "Err")]
)

```

#### Survival Probability

```{r prob-surv-inputs}

# to establish range of exploitation rates used in model; hard-coded for now but
# may allow user input with `params` variable (05-Aug-2019)
mus <- seq(from = 0, to = 0.30, by = 0.01)

# for survival & harvest rates will use defaults unless changed by user
estS <- formals(SurvivalProb)[["estS"]]
estMu <- formals(SurvivalProb)[["estMu"]]
if (!is.na(params[["surv"]])) estS <- params[["surv"]]
if (!is.na(params[["harv"]])) estMu <- params[["harv"]]
# if (is.na(params[["surv"]])) params[["surv"]] <- estS
# if (is.na(params[["harv"]])) params[["harv"]] <- estMu 

# because [["agesHarvest"]] is only length 2 (i.e., range)
agesMu <- params[["agesHarvest"]][1]:params[["agesHarvest"]][2]

```

```{r prob-surv}

# for use in simulations (next steps)
prob_surv <- SurvivalProb(
  ages = prob_survival[["age"]],
  sRate = prob_survival[1:3, "prob"],
  sRateErr = prob_survival[1:3, "se"],
  mu = mus,
  agesMu = agesMu,
  estS = estS,
  estMu = estMu,
  methodSB = TRUE
)

```

## Simulations

Number of simulations: `r params[["sims"]]`.

```{r sims-prob-spawn, eval=TRUE}

sims_prob_spawn <- Simulations(
  data  = prob_spawn2,
  prob  = Prob,
  std   = Err,
  iters = params[["sims"]],
  type  = "spawning"
)

```

```{r sims-num-eggs, eval=TRUE}

sims_num_eggs <- Simulations(
  data  = num_eggs,
  prob  = Count,
  std   = Err,
  iters = params[["sims"]],
  type  = "numeggs"
)

```

```{r sims-prob-surv, eval=TRUE}

sims_prob_surv <- mapply(
  FUN = Simulations,
  prob_surv,
  MoreArgs = list(
    prob  = "Prob",
    std   = "Err",
    recruitment = params[["recruit"]],
    iters = params[["sims"]],
    type  = "survival"
  ),
  SIMPLIFY = FALSE
)

```

```{r sims-fecund}

# either or; should be the same ********
# age_distribution[["FracFemale"]]
# params[["fracFem"]]
# **************************************

sims_fecund <- sims_num_eggs * sims_prob_spawn * params[["fracFem"]]

# hard-coding indices susceptible to problems if data change
final_age <- lapply(prob_surv, FUN = "[", 20, 2:3)

```

## Population Growth Rate

```{r pop-proj}

mu_levels <- setNames(
  object = names(sims_prob_surv),
  nm = names(sims_prob_surv)
)

pop_proj <- lapply(mu_levels, FUN = function(x) {
  PopProjections(
    fSims = sims_fecund,
    sSims = sims_prob_surv[[x]],
    mn = final_age[[x]][["Prob"]],
    sdev = final_age[[x]][["Err"]],
    ageFreq = age_dist2[["Freq"]],
    # ageFreq = age_dist[["freq"]],
    period = 20
    # period = 40
  )
})

pp_names <- dimnames(pop_proj[["mu_0.01"]])[[1]]
pp_length <- length(pop_proj)
pchanges_n <- unique(
  vapply(
    pop_proj[["mu_0.01"]]["pop.changes", ],
    FUN = length,
    FUN.VALUE = numeric(1L)
  )
)

```

`pop_proj` length = `r pp_length`

each element in `pop_proj` contains `r paste0(pp_names, collapse = "; ")`

`pop.changes` in `pop_proj` has `r pchanges_n` values

Using `mu` = 0.01 as an example, we take the mean log of `pop.changes`. The n is given by `params[["sims"]]` or `r params[["sims"]]`

```{r pop-proj-explained}

mll <- vapply(pop_proj[["mu_0.01"]]["pop.changes", ], FUN = function(x) {
  mean(log(x))
}, FUN.VALUE = numeric(1L))

plot(exp(mll), las = 1)
mtext(text = sprintf("n=%s", length(mll)), adj = 0)
abline(h = exp(mean(mll)), col = "orange2")
abline(h = exp(quantile(mll, probs = c(0.025, 0.975))), col = "grey40")
abline(
  h = c(
    exp(mean(mll)) - (sqrt(exp(var(mll)) / length(mll))) * 1.96,
    exp(mean(mll)) + (sqrt(exp(var(mll)) / length(mll))) * 1.96
  ),
  col = c(3, 4)
)

```

```{r lambda-mu}

lambda_mu <- lapply(mu_levels, FUN = function(x) {
  mu <- as.numeric(sub(pattern = "mu_", replacement = "", x = x))
  out <- Lambda(popChanges = pop_proj[[x]]["pop.changes", ])
  out[["MuLevel"]] <- mu
  out
})

lambda_mu <- do.call(what = rbind, args = lambda_mu)
rownames(lambda_mu) <- NULL

```

```{r lambda-bounds}

# x & y values for drawing polygon as lower & upper bounds
poly_list = list(
  x = c(
    lambda_mu[["MuLevel"]][1],
    lambda_mu[["MuLevel"]],
    rev(lambda_mu[["MuLevel"]][-1])
  ),
  y = c(
    lambda_mu[["QuantLow"]][1],
    lambda_mu[["QuantUpp"]],
    rev(lambda_mu[["QuantLow"]][-1])
  )
)

```

```{r calc-ci}

ci <- with(data = lambda_mu, expr = {
  
  se <- sqrt(VarLambda / NumSims)
  
  # where 0.05 is alpha
  se * qnorm(1 - 0.05 / 2)
  
})

```

```{r plot-lambda}

# xlab = expression(bold(paste('Exploitation Rate ( ',mu,' )'))),
#   ylab = expression(bold(paste('Growth Rate ( ', lambda,' )'))),
#   main = expression(bold(paste('Growth Rate (',lambda,') vs Exploitation Rate (',mu,')')))

# create the plot with appropriate limits
plot(
  x = range(lambda_mu[["MuLevel"]]),
  y = range(lambda_mu[, c("QuantLow", "QuantUpp")]),
  type = "n",
  panel.last = abline(h = 1, col = "darkorange", lty = 1, lwd = 0.25),
  panel.first = polygon(poly_list, col = "grey90", border = NA)
)

# add data (mean lambda over mu)
lines(
  x = lambda_mu[["MuLevel"]],
  y = lambda_mu[["MeanLambda"]],
  col = 'grey40',
  lty = 1,
  lwd = 2
)

lines(
  x = lambda_mu[["MuLevel"]],
  y = lambda_mu[["MeanLambda"]] + ci,
  col = 'red',
  lty = 1,
  lwd = 2
)

lines(
  x = lambda_mu[["MuLevel"]],
  y = lambda_mu[["MeanLambda"]] - ci,
  col = 'blue',
  lty = 1,
  lwd = 2
)



```

Parameters list

```{r params-display}

# params[[""]]

Map(function(x, nm) {
  if (all(!is.na(x))) return(x)
  
  # the 3 NAs are estS, estMu, & age0
  if (nm %in% "surv") return(estS)
  if (nm %in% "harv") return(estMu)
  
  age0
  
}, params, names(params))

```

