---
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_knit$set(
  root.dir = "~/RProjects/SturgeonPopMetrics/",
  global.par = TRUE
)

```

```{r global-par}

par(
  mar = c(2.75, 2.5, 1.5, 0.5),
  oma = c(0.5, 0.5, 0.5, 0.5),
  mgp = c(1.75, 0.5, 0)
)

```

```{r load-data}

# SB's data from 'size_selective_data.xlsx', tab 'Fl'; changing column names to
# remove inported prefixed 'X', though not sure it matters for this excercise
sbdata <- read.csv(file = "model/SBData.csv", header = TRUE)
colnames(sbdata) <- c("FL", 6:8)

```

```{r variables}

# could use either; S. Blackburn rounded to one decimal
# msize <- c(6, 7, 8) * 2.54
msize <- c(15.2, 17.8, 20.3)

# represents fishing power of net by mesh size (8" mesh fished at twice the
# effort as typical net config is 8"-6"-7"-8")
rpwr <- c(1, 1, 2)

```

## Starting Values

starting values for net fit model

```{r starting-values}

# SB's starting values (X0) for gear selectivity model; origin unknown & I am
# not sure these values make sense based on length-frequency modes; starting
# values I thought should be based somewhat on existing data; find these values
# in file 'SELECT.R' from SB
X00 <- c(40, 6, 10, 10, 1)

# using functions developed in `spopmodel` to get starting values for comparison
# to SB's; here we select values for the 'binorm.sca' model
X01 <- spopmodel:::GetX0(
  svals = spopmodel:::StartVals(
    len = spopmodel::trammel_catch[["FL"]],
    mesh = spopmodel::trammel_catch[["MeshSize"]]
  )
)[["binorm.sca"]]

# could try other starting values: somewhat arbitrary based on lf distribution
# X01 <- c(75, 15.5, 140, 20.5, 0.62)

```

Display of starting values generated from average of modes by mesh size. Values are markedly different from those used by SB.

```{r display-sbX01}

X01

```

## Model Fit

Note: SB used default for net fishing strength, which assumes equal effort.

```{r model-fit}

# use Millar's NetFit function to apply 'binorm.sca' model to data; SB chose
# 'binorm.sca' for the sturgeon model; gear selectivity model with lowest
# deviance is often the one chosen
fit0 <- spopmodel:::NetFit(
  Data = sbdata,
  Meshsize = msize,
  x0 = X00, 
  rtype = "binorm.sca"
)

# same model but now using different starting values + relative power of 1, 1, 2
fit1 <- spopmodel:::NetFit(
  Data = sbdata,
  Meshsize = msize,
  x0 = X01, 
  rtype = "binorm.sca", 
  rel.power = rpwr
)

# interestingly, below does not work when including un-equal effort, well it
# works but results are head scratching (e.g., large negative deviance)
fit0_rpwr <- spopmodel:::NetFit(
  Data = sbdata,
  Meshsize = msize,
  x0 = X00, 
  rtype = "binorm.sca", 
  rel.power = rpwr
)

```

```{r rel-retention}

# get relative retention from model fit; will standardize in next chunk
rr0 <- spopmodel::RelativeRetention(fit0, standardize = FALSE)
rr1 <- spopmodel::RelativeRetention(fit1, standardize = FALSE)

```

```{r rr-standardized}

# for standardization summing relative retention by length bin & then dividing
# by max value (method used by SB)
rr0_tot <- rowSums(rr0[["Data"]][2:4])
rr1_tot <- rowSums(rr1[["Data"]][2:4])

# to compare relative retention values using different arguments supplied to
# NetFit()
rr0_stand <- rr0_tot / max(rr0_tot)
rr1_stand <- rr1_tot / max(rr1_tot)

```

```{r plot-rel-retention}

# for x-values commonly used for line plotting
xx <- rr0[["Data"]][["FL"]]

# for line size & colors
lwd <- 4L
col0 <- adjustcolor(col = "steelblue", alpha.f = 0.5)
col1 <- adjustcolor(col = "orange2", alpha.f = 0.5)

plot(
  x = xx,
  y = seq(from = 0, to = 1, length.out = 34),
  type = "n",
  xlab = "FL bin",
  ylab = "Relative retention"
)

lines(
  x = xx,
  y = rr0_stand,
  col = col0,
  lwd = lwd
)

lines(
  x = xx,
  y = rr1_stand,
  col = col1,
  lwd = lwd
)

legend(
  x = max(xx),
  y = 1.15,
  legend = c("SB", "Alt"),
  col = c(col0, col1), 
  lwd = lwd,
  xpd = TRUE,
  xjust = 1, 
  ncol = 2,
  bty = "n"
)

```

Now adjust catch using relative retention values, and then plot. Grey line is raw length frequency distribution (i.e., no gear selectivity applied).

```{r plot-adj-catch}

colNoAdj <- adjustcolor(col = "grey50", alpha.f = 0.5)

plot(
  x = range(xx),
  y = c(0, 212),
  type = "n",
  xlab = "FL bin", 
  ylab = "Count"
)

lines(
  x = xx,
  y = rowSums(rr0[["Freq"]] / rr0_stand),
  col = col0,
  lwd = lwd
)

lines(
  x = xx,
  y = rowSums(rr1[["Freq"]] / rr1_stand),
  col = col1,
  lwd = lwd
)

lines(
  x = xx,
  y = rowSums(sbdata[2:4]),
  col = colNoAdj,
  lwd = lwd
)

legend(
  x = max(xx),
  y = 245,
  legend = c("SB", "Alt", "NoAdj"),
  col = c(col0, col1, colNoAdj), 
  lwd = lwd,
  xpd = TRUE,
  xjust = 1, 
  ncol = 3,
  bty = "n"
)

```

---
CDFW Sportfish  
`r Sys.Date()`
