---
output:
   github_document:
     pandoc_args: --webtex
---

```{r setup, include=FALSE}

knitr::opts_knit$set(
  root.dir = "~/RProjects/SturgeonPopMetrics/",
  global.par = TRUE
)

```

```{r src-file, echo=FALSE}

source(file = "presentations/.base-par.R")

```

## Introduction



## Libraries

We load the `sportfish` package, currently available on GitHub. For now (05-Feb-2019), this is the only package required.

```{r load-libraries}

library(sportfish)
# library(package)

```

## Load Data

We load all `.rds` files from directory `data/baystudy`. To keep our workspace clean, we load these files into a new environment called `BayStudy`.


```{r load-data}

# the data directory for bay study
data_dir <- "data/baystudy"

# list.files(path = data_dir, pattern = ".rds")
BayStudy <- new.env()

ReadRDSFiles(fileDir = data_dir, envir = BayStudy)

# clean up
rm(data_dir)

```

## Variables

Here we create some variables we'll use throughout this process. We create them here and now for convenience.

```{r variables}

# # columns in data by which to aggregate
# # cols_agg <- c("Year", "Survey", "Station", "Net", "Tow", "Species", "AgeCat")
# cols_agg <- c("Year", "Survey", "Bay", "Net", "Series")
# 
# cols_match <- intersect(
#   x = colnames(BayStudy$BayStudyTows),
#   y = colnames(BayStudy$StuCatch)
# )
# 
# ntows <- nrow(BayStudy$BayStudyTows)

```

## Sampling

```{r station-count-all}

station_count_all <- aggregate(
  formula = Station ~ Year + Survey,
  data = BayStudy[["BayStudyTows"]],
  FUN = function(x) length(unique(x))#,
  # subset = Series %in% 1 & Net %in% 2
  # subset = Series %in% 1
  # subset = !Flag & Series %in% 1 & Net %in% 2
)

```

```{r}

with(data = station_count_all, expr = {
  
  xrng <- range(Survey)
  yrng <- range(Year)
  
  rect_adj <- 0.50
  
  clr <- rgb(
    red = 0.3,
    green = 0,
    blue = 0,
    alpha = Station / max(Station)
  )
  
  plot(
    x = xrng,
    y = yrng,
    type = "n",
    bty = "n",
    xaxt = "n",
    yaxt = "n",
    xlab = NA,
    ylab = NA
  )
  
  rect(
    xleft = Survey - rect_adj,
    ybottom = Year - rect_adj,
    xright = Survey + rect_adj,
    ytop = Year + rect_adj,
    col = clr,
    border = NA,
    xpd = FALSE
  )
  
  axis(
    side = 3,
    at = 1:12,
    labels = 1:12,
    col = "transparent",
    col.ticks = "grey30"
  )
  
  axis(
    side = 2,
    at = 1980:2018,
    labels = 1980:2018,
    col = "transparent",
    col.ticks = "grey30",
    las = 1
  )
  
  
})




```


## Save Summary Data

```{r save-yci}
# write.csv(yci, file = "indices/yci_bs.csv", row.names = FALSE)
```

---
CDFW, SportFish Unit  
`r Sys.Date()`
