---
title: "White Sturgeon Age-Length Key: A Comparison of Datasets"
author: "CA Department of Fish and Wildlife"
date: "May 9, 2016"
output:
  html_document: default
  github_document:
    html_preview: true 
    toc: true
    fig_height: 6
    fig_width: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```

```{r File Paths and Libraries}

# file paths for absolute reference to source code and data connection
data_dir <- "C:/Data/jdubois/RDataConnections/Sturgeon/"
data_non_db_dir <- "C:/Data/jdubois/RProjects/1_NonDbData"

# creates file path to file containing CDFW age-length data
file_age_len <- paste(data_non_db_dir, "WstAgeLenData.txt", sep = "/")

# load libraries as needed
library(ggplot2)
library(dplyr)
library(knitr)
library(tidyr)
# library(htmlwidgets)
# library(DT)
#library(reshape2)
#library(scales)

```

```{r Data, results='hide', cache=TRUE}

source(file = paste0(data_dir, "OtherData.R"), echo = TRUE)
# source(file = paste0(data_dir, "TaggingData.R"), echo = TRUE)

# loads CDFW age-length data (circa mid 1970s)
old_al_data <- read.table(
  file = file_age_len,
  header = TRUE,
  sep = "\t", 
  # colClasses = c(
  #   "character", "integer", "POSIXct", "double",
  #   "double", rep("integer", 7), "double",
  #   "integer"
  # ),
  col.names = c(
    "ID", "Species", "Date", "FL", "TL", "Sex", "CapMethod",
    "Location", "Age", "YearClass", "Check", "Age_1", "TLen",
    "NewAge"
  ),
  stringsAsFactors = FALSE
)

# re-format date & add year variable for convenience
old_al_data$Date <- as.Date(old_al_data$Date, format = "%m/%d/%y")
old_al_data$Year <- as.numeric(format(old_al_data$Date, "%Y"))

```

```{r Source Code}

# the source directory
source_dir <- "C:/Data/jdubois/RSourcedCode"

# sources needed functions and methods
source(file = paste(source_dir, "methods_len_freq.R", sep = "/"))
source(file = "source_al_key.R")
source(file = paste(source_dir, "methods_age_freq.R", sep = "/"))
# source(file = paste(source_dir, "functions_global_data_subset.R", sep = "/"))
# source(file = paste(source_dir, "functions_global_desc_stats.R", sep = "/"))

# some cleanup
# rm(data_dir, source_dir, data_non_db_dir, file_age_len)
rm(data_dir, data_non_db_dir, file_age_len)

```

### Introduction
Herein we compare extant age-length data (1973-1976, CDFW) with recently-collected and recently-aged age-length data (2012-present, USFWS). Current data collection and ageing is ongoing, so consider this comparison very preliminary.

Further, herein is a comparison of sorts of the extant data to itself. We (CDFW) recently discovered (in electronic form) the age-length data behind our extant age-length key (see WSTALKEY.xls). Recreating the age-length key, we discovered certain (minor) discrepancies. Assuming, of course, we have all the data (and this is a safe assumption +/- a couple data points), the minor discrepancies may be from manual manipulation not (yet) evident in the .xls file or any metadata. In spite of this, we are confident the 1973-1976 data herein are represented in [Kohlhorst et al. (1980)](ftp://ftp.dfg.ca.gov/Adult_Sturgeon_and_Striped_Bass/White%20sturgeon%20age%20and%20growth%201980.pdf).

### General summary: CDFW Data (1973-1976)

* All data are for White Sturgeon  
* Number of records: `r nrow(old_al_data)`  
* Number of columns: `r ncol(old_al_data)`; "Year" field added for convenience  
* In counts per year (summary below), years 1998 and 2014 are likely entry errors  
* Errant data (i.e., data where year > 1976) shown for reference only  
* Data in fields "Age" and "Age_1" are identical  
* Data in fields "TL" and "TLen" are identical

```{r SummaryCDFW}

# display data structure
str(old_al_data)

# count of records per year
table(CountPerYear = old_al_data$Year, useNA = "ifany")

# display records (n=4) where year is a typo (?)
old_al_data[old_al_data$Year > 1976, ]

# fields are the same
# identical(old_al_data$Age, old_al_data$Age_1)
# identical(old_al_data$TL, old_al_data$TLen)

```

### General summary: USFWS Data (2012-ongoing)

* All data are for White Sturgeon  
* Number of records: `r nrow(wst_usfws_age_len)`  
* Number of columns: `r ncol(wst_usfws_age_len)`
* Fish sampled during CDFW mark-recapture study & during sturgeon fishing derby  

```{r SummaryUSFWS}

# display data structure
str(wst_usfws_age_len)

# count of records per year
table(wst_usfws_age_len$Source, useNA = "ifany")

```

### Length Frequency Distribution

Here we look at the length frequency distributions for both data sets. For the purposes of comparison, we used total length (TL, in centimeters) and binned lengths `r paste0("from ", min(wst_alkey$Bins), " to ", max(wst_alkey$Bins), "+", " by 5 cm")`. See overall stats (for TL) in table below.

```{r LFDist}

# establish l-f breaks for WSTALKEY; note this alk uses TL not FL
# wst_alkey_breaks <- c(seq(from = 21, to = 186, by = 5), Inf)
wst_alkey_breaks <- c(wst_alkey$Bins, Inf)

# splitting variables added for convenience of using GetLenFreq() on dataframe
old_al_data$DateSource <- "CDFW"
wst_usfws_age_len$DateSource <- "USFWS"

# create lf dist for 1973-1976 data - using DateSource to "split" data on 
# although DateSource is all CDFW. Using .dataframe method to enable plotting.
lf_cdfw_old <- GetLenFreq(
  old_al_data,
  colX = TL,
  by = "DateSource", 
  breaks = wst_alkey_breaks,
  intBins = FALSE
)

# create lf dist for USFWS data - using DateSource to "split" data on although
# DateSource is all USFWS. Using .dataframe method to enable plotting.
lf_usfws <- GetLenFreq(
  wst_usfws_age_len,
  colX = TotalLength,
  by = "DateSource", 
  breaks = wst_alkey_breaks,
  intBins = FALSE
)

# desc stats for both cdfw and usfws data
kable(
  rbind(
    lf_cdfw_old$DescStats,
    lf_usfws$DescStats
  )[, c(1, 2, 5:8)],
  format = "markdown",
  col.names = c(
    "Source", "Count", "Mean",
    "Median", "Min", "Max" 
  )
)

```

Below, we plot length frequency distributions for CDFW (top) and USFWS (bottom). 

```{r LfDistPlots}

# plot cdfw lf
plot(
  lf = lf_cdfw_old,
  lens = TL,
  fillBar = NULL,
  type = "POT",
  addMed = FALSE,
  addN = TRUE
) +
  theme_stu_al_key

# plot usfws lf
plot(
  lf = lf_usfws,
  lens = TotalLength,
  type = "POT",
  addMed = FALSE,
  addN = TRUE
)+
  theme_stu_al_key

```

### Age Frequency Distribution

```{r AFDist}

# variables used for min, max output and to develop x-axis labeling in plots
# below
age_range_cdfw <- range(old_al_data$Age)
age_range_usfws <- range(wst_usfws_age_len$Age)

max_age <- max(age_range_cdfw[2], age_range_usfws[2])

```

Age data for CDFW ranged `r paste0(age_range_cdfw, collapse = "-")` and for USFWS `r paste0(age_range_usfws, collapse = "-")`. Plots below display age frequency distribution (CDFW - top, USFWS - bottom). Note: different y-axis & USFWS plot begins at age-3.  

```{r AFDistPlots}

# variables for plotting (_x_axis for custom breaks)
age_x_axis_2 <- scale_x_continuous(
  breaks = seq(from = 0, to = max_age, by = 2),
  expand = c(0.01, 0)
)

age_y_axis_2 <- scale_y_continuous(expand = c(0.01, 0))

# plot CDFW age frequency
ggplot(data = old_al_data, mapping = aes(x = Age)) +
  geom_histogram(binwidth = 1, closed = "left") +
  age_x_axis_2 +
  age_y_axis_2 +
  ylab("Count") +
  theme_stu_al_key
  
# plot USFWS age frequency
ggplot(data = wst_usfws_age_len, mapping = aes(x = Age)) +
  geom_histogram(binwidth = 1, closed = "left") +
  age_x_axis_2 +
  age_y_axis_2 +
  ylab("Count") +
  theme_stu_al_key

# using age to develop al key - need to combine all fish > age-22 to age-22
# (this is per design of WSTALKEY)
# old_al_data$Age[old_al_data$Age > 22] <- 22

# clean up
rm(age_x_axis_2, age_y_axis_2)

```

### Mean Length at Age

Presented below are mean lengths at each age (CDFW - top, USFWS - bottom). Along with means (Avg) are standard deviation (SD) and standard error (SE).

```{r MeanLenAtAge}

# not the best way - but works for now...methods_lf has GetDescStats but I want
# to use the one in _global_desc_stats.R, so loading here for this purpose
source(file = paste(source_dir, "functions_global_desc_stats.R", sep = "/"))

# CDFW - create mean length at age output
mean_len_age_cdfw <- old_al_data %>%
  group_by(Age) %>%
  select(Age, TL) %>%
  do(GetDescStats(.$TL))

# USFWS - create mean length at age output
mean_len_age_usfws <- wst_usfws_age_len %>%
  group_by(Age) %>%
  select(Age, TotalLength) %>%
  do(GetDescStats(.$TotalLength))

# CDFW
kable(
  as.data.frame(mean_len_age_cdfw),
  format = "markdown",
  digits = 3
)

# USFWS
kable(
  as.data.frame(mean_len_age_usfws),
  format = "markdown",
  digits = 3
)

# clean up
rm(source_dir)

```

### Growth Curve: von Bertalanffy Growth Curve Applied

[Kohlhorst et al. (1980)](ftp://ftp.dfg.ca.gov/Adult_Sturgeon_and_Striped_Bass/White%20sturgeon%20age%20and%20growth%201980.pdf) applied the von Bertalanffy formula (below) to the 1973-1976 data. Here we repeat the application of this formula to the 1973-1976 data, plus we apply the formula to USFWS data. Note: here we use all ages. Kohlhorst et al. (1980) used ages 0-21. R code included below for reference.

$l_t = L_∞ * (1 - e^{-k * (age - t_0)})$

where:  

$l_t$ = length (cm TL) at a given age  
$L_∞$ = asymptotic length (or length at age infinity)  
$k$ = curvature parameter (how quickly fish gets to $L_∞$)  
$t_0$ = age at which fish has 0 length  

[von Bertalanffy reference](http://www.fao.org/docrep/w5449e/w5449e05.htm)  

```{r vonBert, echo=TRUE}

# Here we apply the von Bertalanffy model to CDFW data and to USFWS data. 
# Summary output assigned to variable for use later in this report and in 
# plotting. VB algorithm fitted using R's nls() function, which requires 
# starting values (see list argument for 'start' parameter). For reference see 
# page 663 of "The R Book" (Crawley 2007)

# The nls() function requires starting values. Values were choses based on
# Kohlhorst et al. 1980 values.

# for simplicity...
# a = L_sub_inf (200)
# b = k         (0.05)
# c = t_sub_0   (0)

# CDFW
mod_vb_cdfw <- nls(
  formula = TLen ~ a * (1 - (exp(-b * (Age - c)))),
  data = old_al_data,
  start = list(a = 200, b = 0.05, c = 0)
)

# USFWS
mod_vb_usfws <- nls(
  formula = TotalLength ~ a * (1 - (exp(-b * (Age - c)))),
  data = wst_usfws_age_len,
  start = list(a = 200, b = 0.05, c = 0)
)

# get summary of each model
mod_vb_cdfw_sum <- summary(mod_vb_cdfw)
mod_vb_usfws_sum <- summary(mod_vb_usfws)

```

Results of the non-linear least squares model `nls()` are given below (CDFW - top, USFWS - bottom). Though displayed as 0, CDFW results are significant to p < 0.001.

```{r VBCoeff Display}

kable(
  mod_vb_cdfw_sum$coefficients,
  format = "markdown",
  format.args = list(scientific = FALSE),
  row.names = TRUE
)

kable(
  mod_vb_usfws_sum$coefficients,
  format = "markdown",
  format.args = list(scientific = FALSE),
  row.names = TRUE
)

```

```{r XaxisAge}

# setting breaks for x-axis for more convenience and more uniformed look
age_x_axis <- scale_x_continuous(
  minor_breaks = seq(from = 0, to = max_age, by = 1),
  expand = c(0.02, 0)
)

```

Here we plot the von Bertalanffy model output. Points are mean total length (cm) at age.

```{r PlotVb}

# set up formula for y-axis in geom_line() - using sprintf() for convenience &
# automation
vb_yval_cdfw <- sprintf(
  "%f * (1 - exp(-%f * (Age - %f)))", 
  mod_vb_cdfw_sum$coefficients["a", "Estimate"],
  mod_vb_cdfw_sum$coefficients["b", "Estimate"],
  mod_vb_cdfw_sum$coefficients["c", "Estimate"]
)

vb_yval_usfws <- sprintf(
  "%f * (1 - exp(-%f * (Age - %f)))", 
  mod_vb_usfws_sum$coefficients["a", "Estimate"],
  mod_vb_usfws_sum$coefficients["b", "Estimate"],
  mod_vb_usfws_sum$coefficients["c", "Estimate"]
)

# plot with mean length at age and fitted vb curve
ggplot() +
  geom_line(
    mapping = aes_string(x = "Age", y = vb_yval_usfws, colour = '"USFWS"'),
    data = wst_usfws_age_len,
    size = 2, alpha = 3/5
  ) +
  geom_line(
    mapping = aes_string(x = "Age", y = vb_yval_cdfw, colour = '"CDFW"'),
    data = old_al_data,
    size = 2, alpha = 3/5
  ) +
  stat_summary(
    mapping = aes(x = Age, y = TotalLength, colour = "USFWS"),
    data = wst_usfws_age_len,
    fun.y = mean, geom = "point", size = 3
  ) +
  stat_summary(
    mapping = aes(x = Age, y = TLen, colour = "CDFW"),
    data = old_al_data,
    fun.y = mean, geom = "point", size = 3
  ) +
  scale_color_manual(
    "",
    values = c(USFWS = "salmon", CDFW = "steelblue")
  ) +
  age_x_axis +
  labs(y = "Mean total length (cm)") +
  theme_stu_al_key

# checking ages 0:100 - not run

# ggplot(data = data.frame(Age = 0:100)) +
#   geom_line(
#     mapping = aes_string(x = "Age", y = vb_yval_cdfw, colour = '"CDFW"'),
#     size = 2, alpha = 3/5
#   ) +
#   geom_line(
#     mapping = aes_string(x = "Age", y = vb_yval_usfws, colour = '"USFWS"'),
#     size = 2, alpha = 3/5
#   )

```

### Other Plots: raw data
Below we plot total length (cm) as a function of age, with loess line added for reference (see blue line). Points are raw data (CDFW - top figure, USFWS - bottom figure) and are shaded according to degree of overlap.

```{r RawDataPlots}

# plots of cdfw age-length data & usfws age-length data; plotting to show raw
# data & trend without fitting data to any one model (see loess line)

ggplot(data = old_al_data, mapping = aes(x = Age, y = TLen)) +
  geom_smooth(se = FALSE, size = 2) +
  geom_point(alpha = 1/5, size = 3) +
  age_x_axis +
  theme_stu_al_key

ggplot(data = wst_usfws_age_len, mapping = aes(x = Age, y = TotalLength)) +
  geom_smooth(se = FALSE, size = 2) +
  geom_point(alpha = 1/5, size = 3) +
  age_x_axis +
  theme_stu_al_key

```

### CDFW 1970s Data: summary count for some variables  
Below are summary outputs for CDFW data (1973-1976). We are in the process of finding metadata for the codes. Likely, though, for "Sex" 1 = male, 2 = female. 

```{r GeneralInfo}

table(CaptureMethod = old_al_data$CapMethod, useNA = "ifany")
table(Location = old_al_data$Location, useNA = "ifany")
table(Sex = old_al_data$Sex, useNA = "ifany")

```

### CDFW 1970s Data: understanding the raw data
Prior to recent endeavors of ageing sturgeon (i.e., collaboration between CDFW & USFWS), we (CDFW) used the age-length key developed from the 1973-1976 data (see WSTALKEY.xls, tab 'A') to age sturgeon. Recently we found the raw data behind this age-length key. However, in trying to recreate the age-length key with this raw data, we discovered some discrepancies. Twenty of the 34 length bins are identical, leaving 14 with (what I'll call) minor deviations. See "ugly" output below for differences by each length bin. Row 3 of $Data is difference between second row (key of newly found age-length data) and first row (extant key). Row 4 is percent changes as calculated below, where *i* denotes each row. If $IsEqual is TRUE, then keys match for that row.

${change} = \frac{new_i - extant_i}{extant_i} * {100}$  

```{r OldALKey}

# create age-length key from "found" electronic age-length key data

# using age to develop al key - need to combine all fish > age-22 to age-22
# (this is per design of WSTALKEY)
old_al_data$Age[old_al_data$Age > 22] <- 22

new_old_alkey <- MakeALKey(
  dat = old_al_data,
  len = TL,
  age = Age,
  lenBreaks = wst_alkey_breaks,
  breakLabs = wst_alkey_breaks[-35]
)

# convert wst_alk_usfws bins to integer for analytics to follow
new_old_alkey$Bins <- as.integer(as.character(new_old_alkey$Bins))

```

```{r DisplayALK, eval=FALSE}

kable(
  wst_alkey,
  format = "markdown",
  digits = 4
)

kable(
  new_old_alkey,
  format = "markdown",
  digits = 4
)

```

```{r Differences, comment=""}

# variable for looping through sapply below
rows <- 1:nrow(wst_alkey)
names(rows) <- paste0("Bin", wst_alkey[, 1])

# to compare each line (bin) of wst_alkey with the newly-created alkey
lst_diff <- sapply(X = rows, FUN = function(x) {
  
  # bind each line of the old and new alkey
  res <- rbind(
    wst_alkey[x, ],
    round(new_old_alkey[x, ], digits = 4)
  )
  
  # calculate difference and percent change for comparison (setting old key as 
  # "first" value); see http://www.percent-change.com/index.php
  difference <- res[2, ] - res[1, ]
  percent_change <- ((res[2, ] - res[1, ]) / res[1, ]) * 100
  
  # combine for convenient output
  res <- rbind(res, difference, percent_change)
  
  # reset row numbers
  rownames(res) <- 1:nrow(res)
  
  # function output
  list(
    Data = res[, -1],
    # Data = res,
    # Diff = difference,
    # PerChange = percent_change,
    IsEqual = all(res[1, ] - res[2, ] == 0)
  )
  
}, simplify = FALSE)

# ideally would yield number equivalent to number of bins (n=34) but only yields
# 20, so 14 bins of newly-created alkey to not "align" with extant WSTALKEY
# sum(
#   sapply(X = rows, FUN = function(x) {
#     lst_diff[[x]]$IsEqual
#   })
# )

lst_diff

```

To examine differences in the age-length keys, we ran fake data through both, and then compared the difference in frequency at age. Difference measured in direction wst_alkey - alkey from length & age data found in 1973-1976 file (n = 1222). Fake length data was randomly generated, but this is fine since we are interested in the differences and not absolute counts at age.

The red line in the plot is set at y = 0, blue lines at 0.025 & -0.025.

```{r TestKeys}

# mean(old_al_data$TL)
# range(old_al_data$TL)
# sd(old_al_data$TL)

# TODO: look at rnorm instead of sample()
# TODO: look at more than just rnorm - as true lf dist is
#       likely not normally distributed
# TODO: create function to help with showing prop_diff on
#       varying sample size

# generate fake length data; Year variable required in GetLenFreq()
tl_fake <- data.frame(
  Year = 2016,
  # TL = sample(21:200, size = 100000, replace = TRUE),
  TL = rnorm(n = 300, mean = 97, sd = 33.1),
  stringsAsFactors = FALSE
)

# using wst length breaks, get length frequency
lf_fake <- GetLenFreq(
  tl_fake,
  colX = TL,
  by = "Year",
  breaks = wst_alkey_breaks,
  intBins = TRUE
)

# get age frequency using the two age-length keys
af_fake_1 <- GetAgeFreq(lfTable = lf_fake$Freq, alk = wst_alkey)
af_fake_2 <- GetAgeFreq(lfTable = lf_fake$Freq, alk = new_old_alkey)

```

```{r PlotTestKeys}

diff_count <- unlist(af_fake_2[, -1] - af_fake_1[, -1])
prop_change <- diff_count / unlist(af_fake_1[, -1])

# plot differences as extant - newly created
plot(
  colnames(af_fake_1[, -1]),
  # unlist(af_fake_1[, -1] - af_fake_2[, -1]),
  prop_change,
  # diff_count,
  pch = 20, xlab = "Age", ylab = "Proportional change"
)
abline(h = 0, col = 2)
abline(h = 0.025, col = 4, lty = 2, lwd = 0.25)
abline(h = -0.025, col = 4, lty = 2, lwd = 0.25)

```

<a href="#top">back to top</a>  
***  
Report ran: `r Sys.time()`  
End of report
