---
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_knit$set(root.dir = "~/RProjects/SturgeonPopMetrics/")
knitr::opts_chunk$set(echo = FALSE)

```

```{r load-libraries}

library(sportfish)
# library(package)

```

```{r load-data, results='hide'}

# the data directory for bay study
data_dir <- "data/tagging"

# list.files(path = data_dir, pattern = ".rds")
Tagging <- new.env()

ReadRDSFiles(fileDir = data_dir, envir = Tagging)

notes <- readLines(file.path(data_dir, "data-log"), n = 2)
notes <- paste0(notes[1], notes[2])

# clean up
rm(data_dir)

```

```{r variables}

# to remove 1955 (1 day sampling) & rescue fish (from 2011 not caught in trammel
# net operations)
b_1955_rescue <- substitute(
  RelYear != 1955 & !(TaggingLoc %in% c("FremontW", "TisdaleBP"))
)

# to get net fathom hours & soak time
effort <- EffortStu(
  data = Tagging[["Effort"]],
  sStart = SetStart,
  sEnd = SetEnd,
  rStart = RetrieveStart,
  rEnd = RetrieveEnd ,
  netLength = Fathoms
)

```

## Project Overview

To display California-based White Sturgeon population metrics. Metrics herein may be used in status reports for said White Sturgeon. Further, to summarize --- and in some cases analyze --- Green Sturgeon data.

## Sampling - Mark-Recapture

```{r annual-date-range}

with(data = Tagging[["Effort"]], expr = {
  
  dates <- as.Date(paste("2019", format(RelDate, format = "%m-%d"), sep = "-"))
  
  b <- dates > as.Date("2019-07-31")
  
  rngx <- range(RelYear[b])
  rngy <- range(dates[b])
  
  # to plot x-axis tick labels every 5 years
  xticks <- (rngx[1]:rngx[2]) %% 5 == 0
  
  plot(
    x = rngx,
    y = rngy,
    type = "n",
    xaxt = "n",
    # yaxt = "n",
    xlim = rngx + c(1.25, -1.25),
    las = 1,
    xlab = "Year",
    ylab = "Sampling dates"
  )
  
  # for custom x-axis ticks & tick labels
  axis(
    side = 1,
    at = (rngx[1]:rngx[2])[xticks],
    labels = (rngx[1]:rngx[2])[xticks],
    col = "transparent",
    col.ticks = "grey30"
  )
  
  par(xaxp = c(rngx, diff(rngx)))
  # par(yaxp = c(18109, 18230, 7))
  grid(lwd = 1000, col = "grey90")
  grid(lty = 1, col = "white", lwd = 1)
  
  # get min & max date by release year
  bounds <- aggregate(
    x = dates[b],
    by = list(Year = RelYear[b]),
    FUN = function(x) c(min(x), max(x))
  )
  
  # for ease of plotting
  bounds <- data.frame(
    Year = bounds[["Year"]],
    Min = as.Date(bounds[["x"]][, 1], origin = "1970-01-01"),
    Max = as.Date(bounds[["x"]][, 2], origin = "1970-01-01")
  )
  
  # all sampling dates (as small points)
  points(
    RelYear[b],
    dates[b],
    pch = 20,
    # pch = ".",
    cex = 0.01,
    col = "grey60"
  )
  # beginning sampling date per year
  points(
    x = bounds[["Year"]],
    y = bounds[["Min"]],
    pch = 20,
    col = "steelblue"
  )
  # ending sampling date per year
  points(
    x = bounds[["Year"]],
    y = bounds[["Max"]],
    pch = 20,
    col = "orange2"
  )
  
  legend(
    x = rngx[2],
    y = as.numeric(rngy[2]) * 1.001,
    legend = c("start date", "end date"),
    pch = 20,
    col = c("steelblue", "orange2"),
    xpd = TRUE,
    xjust = 1,
    ncol = 2,
    bty = "n"
  )

  # output dates not plotted
  # RelDate[!b]
})

```

```{r split-effort}

split_vars <- c("RelDate", "Vessel", "NetSet", "TaggingLoc")

b <- eval(expr = b_1955_rescue, envir = Tagging[["Effort"]])

split_effort_data <- split(
  data.frame(
    Tagging[["Effort"]][b, split_vars],
    NetFH = effort[["Effort"]][b],
    SoakTime = effort[["SoakTime"]][b],
    TimeFished = effort[["TimeFished"]][b]
  ),
  f = Tagging[["Effort"]][b, "RelYear"]
)

# clean up
rm(b)

```

```{r effort-summary}

fn_out <- c(
  "Year", "Days", "Vessels", "NetSets",
  "Locations", "NetFH", "Hours", "MeanTF", "SdTF"
)

fn_out <- setNames(
  object = rep(0, times = length(fn_out)),
  nm = fn_out
)

effort_summary <- vapply(split_effort_data, FUN = function(x) {
  
  nfh <- sum(x[["NetFH"]], na.rm = TRUE)
  if (all(nfh == 0)) nfh <- NA
  
  hours <- sum(x[["TimeFished"]], na.rm = TRUE)
  if (all(hours == 0)) hours <- NA
  
  out <- c(
    unique(as.numeric(format(x[["RelDate"]], format = "%Y"))),
    length(unique(x[["RelDate"]])),
    length(unique(x[["Vessel"]])),
    length(x[["NetSet"]]),
    length(unique(x[["TaggingLoc"]])),
    nfh,
    hours,
    mean(x[["TimeFished"]], na.rm = TRUE) * 60,
    sd(x[["TimeFished"]], na.rm = TRUE) * 60
  )
  
  names(out) <- names(fn_out)
  out
  
}, FUN.VALUE = fn_out)

effort_summary <- data.frame(t(effort_summary), row.names = NULL)

```

```{r plot-effort, fig.height=6, fig.asp=1}

# head(effort_summary)

# for stacked plot
mat_layout <- matrix(data = 1:3, nrow = 3, ncol = 1, byrow = TRUE)
nf <- layout(mat = mat_layout)
# layout.show(n = nf)

# for keeping bottom & left spaces to display axes title
# par(oma = c(4, 5, 1, 1), cex.axis = 1.5, family = "sans")
par(oma = c(4, 4, 0.5, 1), cex.axis = 1.05)

with(data = effort_summary, expr = {
  
  l <- list(Days = Days, NetSets = NetSets, NetFH = NetFH)
  
  rngy <- vapply(l, FUN = function(x) {
    
    x <- Filter(f = Negate(is.na), x)
    
    mn <- min(x)
    mx <- max(x)
    
    c(Low = 0, Min = mn, Max = mx)
    
  }, FUN.VALUE = numeric(3L))
  # end vapply
  
  rngx <- range(Year)
  all_years <- rngx[1]:rngx[2]
  xticks <- all_years %% 5 == 0
  
  lkp_vessel_col <- c(`1` = "steelblue", `2` = "orange2")
  
  n <- 1
  
  lapply(dimnames(rngy)[[2]], FUN = function(y) {
    
    mar3 <- 1.5
    if (n > 1) mar3 <- 0.5
    
    n <<- n + 1
    
    # par(mar = c(0.1, 0.5, 0.5, 0.1), mgp = c(1, 0.6, 0))
    par(mar = c(0, 0.5, mar3, 0.1), mgp = c(1, 0.6, 0))
    # par(mar = c(0.1, 0.5, 1.0, 0.1), mgp = c(1, 0.6, 0))
    
    plot(
      x = rngx,
      y = rngy[c("Low", "Max"), y],
      type = "n",
      xlim = rngx + c(1.5, -1.5),
      xlab = NA,
      ylab = NA,
      yaxt = "n",
      xaxt = "n",
      las = 1
    )
    
    y_format <- AxisFormat(axTicks(side = 2))
    
    axis(
      side = 2,
      at = axTicks(side = 2),
      labels = y_format[["Labels"]],
      col = "transparent",
      col.ticks = "grey30",
      las = 1,
      cex = 1.25
    )
    
    txt <- y_format$AxisTitle(y)
    
    # txt <- if (max(axTicks(side = 2) < 100)) y
    
    mtext(text = txt, side = 2, line = 2)
    
    par(xaxp = c(rngx, diff(rngx)))
    grid(lwd = 1000, col = "grey90")
    grid(lty = 1, col = "white", lwd = 1)
    
    lines(
      x = Year,
      y = get(y),
      type = "h",
      col = lkp_vessel_col[Vessels],
      lwd = 5,
      lend = 1
    )
    
  })
  # end lapply
  # layout(mat = 1)
  
  # print(axTicks(side = 1))
  
  axis(
    side = 1,
    at = all_years[xticks],
    labels = all_years[xticks],
    col = "transparent",
    col.ticks = "grey30",
    cex = 1.25
  )
  
  mtext(text = "Year", side = 1, line = 2)
  
  # 
  # plot(Year, Days, col = Vessels, type = "h")
  # plot(Year, NetSets, col = Vessels, type = "h")
  # plot(Year, NetFH, col = Vessels, type = "h")
  # dimnames(rngy)
  all_years[xticks]
  axTicks(side = 1)
  
  
  layout(mat = 1)
  par(lend = 1)
  
  legend(
    x = rngx[2],
    y = 40000 * 1.1,
    legend = c("One Vessel", "Two Vessels"),
    col = c("steelblue", "orange2"),
    lwd = 5,
    border = NA,
    # col = NA,
    bty = "n",
    ncol = 2,
    xpd = TRUE,
    xjust = 1,
    yjust = 0.5,
    x.intersp = 0.5,
    seg.len = 2
  )
  
})

```

<!--
Here we show annual effort for years we performed mark-recapture fieldwork using trammel nets. Field names are explained below.

- Year: sampling year
- Days: number of calendar days fieldwork performed
- Vessels: number of vessels used in fieldwork
- NetSets: total number of net sets
- Locations: number of locations sampled (large scale, e.g., Suisun Bay)
- NetFH: net-fathom hours (unit of effort used to calculate CPUE)
- Hours: total fishing time in hours
- MeanTF: average annual fishing time (hours)
- SdTF: standard deviation annual fishing time (hours)
-->

```{r effort-display, eval=FALSE}

knitr::kable(effort_summary, format = "pandoc")

```

## Catch

```{r catch}

# aggregate(
#   formula = NetSetId ~ RelYear + Species + StuType,
#   data = Tagging[["Sturgeon"]],
#   FUN = length,
#   subset = eval(b_1955_rescue)
# )

catch <- with(data = Tagging[["Sturgeon"]], expr = {
  # favoring aggregate.data.frame method rather than formula method for use of
  # `drop` parameter - keeps years in order for when no GST were caught
  
  b <- eval(b_1955_rescue)
  
  out <- aggregate(
    StuType[b],
    by = list(Year = RelYear[b], Species = Species[b]),
    FUN = function(x) {
      c(Tagged = sum(x %in% "Tag"), Total = length(x))
    },
    drop = FALSE
  )
  
  # for year (or years) with 0 catch (e.g., 2016 no GST)
  out$x[is.na(out[["x"]])] <- 0
  
  data.frame(
    Year = out[["Year"]],
    Species = out[["Species"]],
    out[["x"]],
    stringsAsFactors = FALSE
  )
  
})

```

Here we show by species (Green Strugeon and White Sturgeon) annual counts tagged and total caught (including recaptures from within season or previous seasons & for 2007 & 2019 includes PIT tags) during trammel net operations. In 2013, we stopped disc tagging Green Sturgeon. (Eventually, we'll summarize tagging separately by PIT & disc.)

```{r plot-catch}

# for stacked plot
mat_layout <- matrix(data = 1:2, nrow = 2, ncol = 1, byrow = TRUE)
nf <- layout(mat = mat_layout)
# layout.show(n = nf)

# for keeping bottom & left spaces to display axes title
# par(oma = c(4, 5, 1, 1), cex.axis = 1.5, family = "sans")
par(oma = c(3, 3, 0.5, 1), cex.axis = 1.05)

with(data = catch, expr = {
  
  # to create empty plot
  rngx <- range(Year)
  # rngy <- range(Total)
  
  # for x-axis labels every 5 years
  all_years <- rngx[1]:rngx[2]
  xtick <- all_years %% 5 == 0
  # print(xtick)
  
  # for plotting both gear species
  vals <- lapply(list("Green", "White"), FUN = function(g) {
    
    par(mar = c(0.1, 0.5, 1.0, 0.1), mgp = c(1, 0.6, 0))
    
    # to get data specific to gear type
    b <- Species %in% g & Total != 0
    
    # needed here due to the vast difference in catch by species
    rngy <- range(Total[b])
    
    # to plot annual ticks every 5 years
    # xtick <- all_years %% 5 == 0
    
    plot(
      x = rngx,
      y = c(min(0, rngy[1]), rngy[2]),
      type = "n",
      xaxt = "n",
      las = 1,
      yaxt = "n",
      # xlab = "Year",
      # ylab = sprintf(fmt = "Days sampled (%s)", g),
      xlim = rngx + c(1, -1)
    )
    
    # print(axTicks(side = 2))
    y_format <- AxisFormat(axTicks(side = 2))

    # custom y-axis tick display
    axis(
      side = 2,
      at = axTicks(side = 2),
      labels = y_format[["Labels"]],
      # labels = axTicks(side = 2) / 1000,
      col = "transparent",
      col.ticks = "grey30",
      cex = 1.05,
      las = 1
    )
  
    # for grey background & white gridlines
    par(xaxp = c(rngx, diff(rngx)))
    grid(lwd = 1000, col = "grey90")
    grid(lty = 1, col = "white", lwd = 1)
  
    # plot total data as bar
    lines(
      # x = X1[b] - 0.2,
      x = Year[b],
      y = Total[b],
      lend = 1,
      lwd = 7,
      # col = "orange2",
      # col = rgb(0, 0, 0, 0.75),
      col = "grey20",
      type = "h"
    )
    
    # plot tagged data as bar
    lines(
      # x = X1[b] - 0.2,
      x = Year[b],
      y = Tagged[b],
      lend = 1,
      lwd = 2,
      # col = "orange2",
      # col = rgb(0, 0, 0, 0.75),
      col = "orange2",
      type = "h"
    )
    
    # for identifying which plot is which
    mtext(
      text = sprintf(fmt = "%s Sturgeon", g),
      side = 3,
      line = 0,
      adj = 0
    )
    
    # x-axis labeling (different for greens & whites)
    mtext(text = y_format$AxisTitle("Count"), side = 2, line = 2)
    
    # output for convenience of custom x-axis
    # xtick
    # sum(b)
    y_format
  })
  # end lapply()
  
  # resent to one plot
  layout(mat = 1)
  # print(vals)
  
  # custom x-axis
  axis(
    side = 1,
    # at = RelYear[vals[[1]]],
    # labels = RelYear[vals[[1]]],
    at = all_years[xtick],
    labels = all_years[xtick],
    col = "transparent",
    col.ticks = "grey30",
    cex = 1.05
  )
  
  # axis labels
  mtext(text = "Year", side = 1, line = 1.75)
  
  par(lend = 1)
  
  legend(
    x = rngx[2],
    y = max(Total) * 1.08,
    legend = c("Total", "Tagged"),
    col = c("grey20", "orange2"),
    lwd = c(10, 2),
    border = NA,
    # col = NA,
    bty = "n",
    ncol = 2,
    xpd = TRUE,
    xjust = 1,
    yjust = 0,
    x.intersp = 0.5,
    seg.len = 2
  )
  
})

```

```{r catch-display, eval=FALSE}

knitr::kable(
  reshape(
    data = catch,
    timevar = "Species",
    idvar = "Year",
    direction = "wide",
    sep = " "
  ),
  format = "pandoc"
)

```

## Tag Returns

```{r eval=FALSE}


with(data = Tagging[["TagReturns"]], expr = {
  
  recy <- as.numeric(format(DateCaptured2, format = "%Y"))
  
  table(RelYear, recy, useNA = "ifany")
  
  table(recy, useNA = "ifany")
  
  plot(RelYear, recy, type = "n")
  
  text(RelYear, recy, labels = try$RetYear)
  
  
})

try <- TagRetYear(
  data = Tagging[["TagReturns"]],
  dcap = DateCaptured2,
  drel = RelDate
)

plot(Tagging$TagReturns$RelYear, try$DAL, col = rgb(0, 0, 0, 0.1), pch = 20)


hist(try$RetYear)
hist(try$DAL)

```

```{r eval=FALSE}

# closer to what I was thinking

with(data = Tagging, expr = {
  
  tRY <- TagRetYear(
    data = TagReturns,
    dcap = DateCaptured2,
    drel = RelDate
  )
  
  
  tbl <- table(
    tRY[["RetYear"]],
    TagReturns[["RelYear"]],
    useNA = "ifany",
    dnn = c("Tagging Year", "Return Year")
  )
  
  plot(x = tRY[["RetYear"]], y = TagReturns[["RelYear"]],
       # col = hsv(alpha = tbl / max(tbl)))
       col = heat.colors(n = 5, alpha = tbl / max(tbl)))
  # text(x = tRY[["RetYear"]], y = TagReturns[["RelYear"]], labels = tbl)
  
  tbl / max(tbl)
  
})


```



---
Contact [Jason DuBois](mailto:jason.dubois@wildlife.ca.gov)  
CDFW, SportFish Unit  
**Note**: `r notes`  
