---
title: |
  | Year-class Indices & Other Population Metrics
  | Sturgeon Status and Review:
author: "Andrew Danos, Jason DuBois, Randy Baxter"
date: "March 7, 2019"
output:
  slidy_presentation:
    duration: 15
    footer: CDFW 2019
    keep_md: no
  ioslides_presentation: default
  powerpoint_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(
  root.dir = "~/RProjects/SturgeonPopMetrics/",
  global.par = TRUE
)

```

```{r load-data, include=FALSE}

# YCI (bay study & salvage)
yci_bs <- read.csv(file = "indices/yci_bs.csv")
yci_salvage <- read.csv(file = "indices/yci_salvage.csv")

# Card (summary, length summary, & angler catch raw data)
card_summary <- read.csv(file = "fishery/card_summary.csv")
card_summary <- subset(card_summary, subset = Year < 2019)

card_sp_length <- read.csv(
  file = "fishery/card_sp_length.csv",
  stringsAsFactors = FALSE
)
card_sp_length <- subset(card_sp_length, subset = Year < 2019)

AnglerCatch <- readRDS("data/card/AnglerCatch.rds")

# CPFV CPUE

SFESuccessful <- readRDS("data/cpfv/SFESuccessful.rds")

# Mark-Recapture (all sturgeon, annual cpue summary)
Sturgeon <- readRDS("data/tagging/Sturgeon.rds")
Sturgeon <- subset(Sturgeon, subset = RelYear > 1967)

CpueWST <- read.csv(file = "abundance/cpue_wst.csv")
tagging_years <- CpueWST[["year"]] # for dataset plot
CpueWST <- subset(CpueWST, subset = year > 1967)

WSTHarvest <- readRDS("data/tagging/WSTHarvest.rds")
WSTHarvest <- subset(WSTHarvest, subset = RelYear > 1997)

# Blackburn model sims
lambda_mu <- read.csv(file = "presentations/IEP/BlackburnModelSims.csv")

```

```{r source-files, include=FALSE}

source(file = "presentations/.base-par.R")

```

```{r libraries, include=FALSE}

library(sportfish)

```

# Datasets

```{r data-sets}

# (1) baystudy    1980-present
# (2) cpfv        log books 1936-1940, 1947-present
# (3) reportcard  2007-present
# (4) salvage     1957-present (CVP),; 1968-present (SWP)
# (5) tagging     intermittently since 1954
# (6) nativefish  ??

years_ds <- 1936:2018
evens <- years_ds %% 2 == 0

# set up plot
plot(
  x = range(years_ds),
  y = c(1, 5),
  type = "n",
  xaxt = "n",
  yaxt = "n",
  # xlab = "Year",
  xlab = NA,
  # ylab = "Dataset"
  ylab = NA
)

axis(side = 1, at = years_ds[evens], labels = years_ds[evens], tck = -0.02)
axis(side = 1, at = years_ds, labels = NA, tck = -0.01)
axis(
  side = 2,
  at = 5:1,
  labels = c("BayS", "CPFV", "Card", "Slvg", "MrkR"),
  las = 1
)

# for ease of changing points and point size
pt <- "|"
ps <- 2
coln <- "orange3"    # color for data not used (n)
colu <- "royalblue3" # color for data used (u)

# bay study
points(1980:2018, rep(5, times = 39), pch = pt, cex = ps, col = colu)
# cpfv
points(1980:2018, rep(4, times = 39), pch = pt, cex = ps, col = colu)
points(1947:1979, rep(4, times = 33), pch = pt, cex = ps, col = coln)
points(1936:1940, rep(4, times = 5), pch = pt, cex = ps, col = coln)
# card
points(2007:2018, rep(3, times = 12), pch = pt, cex = ps, col = colu)
# salvage
points(1957:1979, rep(2, times = 23), pch = pt, cex = ps, col = coln)
points(1980:2018, rep(2, times = 39), pch = pt, cex = ps, col = colu)
# tagging
points(tagging_years[1:3], rep(1, times = 3), pch = pt, cex = ps, col = coln)
points(
  tagging_years[4:length(tagging_years)],
  rep(1, times = length(tagging_years) - 3),
  pch = pt,
  cex = ps,
  col = colu
)

# red point for year of no tagging (2018)
points(2018, 1, pch = pt, cex = ps, col = 2)

# clean up
rm(evens, years_ds)

```

# Bay Study YCI - White Sturgeon

```{r plot-bs_wyci}

years_bs <- yci_bs[["Year"]]
evens <- years_bs %% 2 == 0
div <- 100
range_wyci <- range(yci_bs[["WYCI"]], na.rm = TRUE)

# plot(WYCI ~ Year, data = yci_bs, type = "h", lend = 2)
# plot(WYCI ~ Year, data = yci_bs, type = "b")

plot(
  x = range(years_bs),
  y = range_wyci,
  type = "n",
  xaxt = "n",
  yaxt = "n",
  # xlab = "Year",
  xlab = "",
  ylab = sprintf("Year-class index (x %s)", div)
)
yvals <- axTicks(side = 2)

lines(x = years_bs, y = yci_bs[["WYCI"]], lty = 2, lwd = 0.4, col = "grey40")
# points(
#   x = years_bs,
#   y = yci_bs[["WYCI"]],
#   cex = 1.5,
#   pch = 21,
#   col = "white",
#   bg = "white"
# )
points(
  x = years_bs,
  y = yci_bs[["WYCI"]],
  cex = 2,
  pch = 21,
  col = "white",
  bg = "royalblue3"
)
axis(side = 1, at = years_bs[evens], labels = years_bs[evens], tck = -0.02)
axis(side = 1, at = years_bs, labels = NA, tck = -0.01)
axis(side = 2, at = yvals, labels = yvals / div, las = 1)
mtext(
  text = "Reference: Fish 2010",
  side = 3,
  line = 0.25,
  adj = 0,
  col = "grey40"
)

```

# Bay Study YCI - Green Sturgeon

```{r plot-bs_gyci}

range_gyci <- range(yci_bs[["GYCI"]], na.rm = TRUE)
div <- 10

plot(
  x = range(years_bs),
  y = range_gyci,
  type = "n",
  xaxt = "n",
  yaxt = "n",
  # xlab = "Year",
  xlab = "",
  ylab = sprintf("Year-class index (x %s)", div)
)
yvals <- axTicks(side = 2)

lines(x = years_bs, y = yci_bs[["GYCI"]], lty = 2, lwd = 0.4, col = "grey40")
# points(
#   x = years_bs,
#   y = yci_bs[["GYCI"]],
#   cex = 1.5,
#   pch = 21,
#   col = "white",
#   bg = "white"
# )
points(
  x = years_bs,
  y = yci_bs[["GYCI"]],
  cex = 2,
  pch = 21,
  col = "white",
  bg = "royalblue3"
)
axis(side = 1, at = years_bs[evens], labels = years_bs[evens], tck = -0.02)
axis(side = 1, at = years_bs, labels = NA, tck = -0.01)
axis(side = 2, at = yvals, labels = yvals / div, las = 1)
mtext(
  text = "Reference: Heublein et al. (in press)",
  side = 3,
  line = 0.25,
  adj = 0,
  col = "grey40"
)

# yci_bs[["GYCI"]]

```

# Salvage YCI - White Sturgeon

```{r plot-salwyci}

sal_yr <- range(yci_salvage[["Year"]])
sal_wyci <- range(yci_salvage[["WYCI"]])

sal_yr_labels <- seq(from = sal_yr[1], to = sal_yr[2], by = 2)

ly <- layout(
  mat = matrix(data = c(1, 2), nrow = 2, ncol = 1, byrow = TRUE),
  heights = c(5, 6)
)

# layout.show(ly)

op <- par(mar = c(0.15, 4.5, 1, 1), las = 1)
plot(
  x = sal_yr,
  y = sal_wyci,
  type = "n",
  xaxt = "n",
  xlab = NA,
  ylab = "YCI-SWP"
)

mtext(
  text = "Reference: Heublein et al. (in press)",
  side = 3,
  line = 0,
  adj = 0,
  col = "grey40"
)

lines(
  formula = WYCI ~ Year,
  data = yci_salvage,
  subset = Facility == 1,
  cex = 1.5,
  lwd = 0.75,
  col = "grey40",
  lty = 2
)
points(
  formula = WYCI ~ Year,
  data = yci_salvage,
  subset = Facility == 1,
  cex = 1.5,
  pch = 21,
  col = "white",
  bg = "royalblue3"
)

par(op)
op <- par(mar = c(4, 4.5, 0.15, 1), las = 1)

plot(
  x = sal_yr,
  y = sal_wyci,
  type = "n",
  xaxt = "n",
  xlab = NA,
  ylab = "YCI-CVP"
)
lines(
  formula = WYCI ~ Year,
  data = yci_salvage,
  subset = Facility == 2,
  cex = 1.5,
  lwd = 0.75,
  col = "grey40",
  lty = 2
)
points(
  formula = WYCI ~ Year,
  data = yci_salvage,
  subset = Facility == 2,
  cex = 1.5,
  pch = 21,
  col = "white",
  bg = "royalblue3"
)

axis(side = 1, at = sal_yr_labels, labels = sal_yr_labels, tck = -0.02)
axis(side = 1, at = sal_yr[1]:sal_yr[2], labels = NA, tck = -0.01)
par(op)

```

# Salvage YCI - Green Sturgeon

```{r plot-salgyci}

sal_gyci <- range(yci_salvage[["GYCI"]])

ly <- layout(
  mat = matrix(data = c(1, 2), nrow = 2, ncol = 1, byrow = TRUE),
  heights = c(5, 6)
)

op <- par(mar = c(0.15, 4.5, 1, 1), las = 1)
plot(
  x = sal_yr,
  y = sal_gyci,
  type = "n",
  xaxt = "n",
  xlab = NA,
  ylab = "YCI-SWP"
)

mtext(
  text = "Reference: Heublein et al. (in press)",
  side = 3,
  line = 0,
  adj = 0,
  col = "grey40"
)

lines(
  formula = GYCI ~ Year,
  data = yci_salvage,
  subset = Facility == 1,
  cex = 1.5,
  lwd = 0.75,
  col = "grey40",
  lty = 2
)
points(
  formula = GYCI ~ Year,
  data = yci_salvage,
  subset = Facility == 1,
  cex = 1.5,
  pch = 21,
  col = "white",
  bg = "royalblue3"
)

par(op)
op <- par(mar = c(4, 4.5, 0.15, 1), las = 1)

plot(
  x = sal_yr,
  y = sal_gyci,
  type = "n",
  xaxt = "n",
  xlab = NA,
  ylab = "YCI-CVP"
)
lines(
  formula = GYCI ~ Year,
  data = yci_salvage,
  subset = Facility == 2,
  cex = 1.5,
  lwd = 0.75,
  col = "grey40",
  lty = 2
)
points(
  formula = GYCI ~ Year,
  data = yci_salvage,
  subset = Facility == 2,
  cex = 1.5,
  pch = 21,
  col = "white",
  bg = "royalblue3"
)

axis(side = 1, at = sal_yr_labels, labels = sal_yr_labels, tck = -0.02)
axis(side = 1, at = sal_yr[1]:sal_yr[2], labels = NA, tck = -0.01)
par(op)

```

# YCI Correlation (White Sturgeon)

```{r yci-corr}

layout(1)

years <- 1980:2016

b_bst <- yci_bs[["Year"]] %in% years
b_sal <- yci_salvage[["Year"]] %in% years
b_swp <- yci_salvage[["Facility"]] %in% 1

mat <- matrix(
  data = c(
    yci_bs[b_bst, "WYCI"],
    yci_salvage[b_sal & b_swp, "WYCI"],
    yci_salvage[b_sal & !b_swp, "WYCI"],
    yci_bs[b_bst, "GYCI"],
    yci_salvage[b_sal & b_swp, "GYCI"],
    yci_salvage[b_sal & !b_swp, "GYCI"]
  ),
  nrow = length(years),
  dimnames = list(
    NULL,
    c("WBST", "WSWP", "WCVP", "GBST", "GSWP", "GCVP")
  )
)

# cor_res <- Hmisc::rcorr(mat)
# cor_res[["r"]]

```

```{r wst-cor}

pairs(
  mat[, 1:3],
  upper.panel = PanelCorr(out = "text"),
  lower.panel = PanelCorr()
)
mtext(
  text = "Reference: Heublein et al. (in press)",
  side = 3,
  line = 0,
  adj = 0,
  col = "grey40"
)

```

# YCI Correlation (Green Sturgeon)

```{r gst-cor}

pairs(
  mat[, 4:6],
  upper.panel = PanelCorr(out = "text"),
  lower.panel = PanelCorr()
)
mtext(
  text = "Reference: Heublein et al. (in press)",
  side = 3,
  line = 0,
  adj = 0,
  col = "grey40"
)

```

# Report Card - White Sturgeon

```{r plot-kept_rel}

div <- 1000
range_yr <- range(card_summary[["Year"]])
years <- card_summary[["Year"]]

ymax <- max(card_summary[, c("WSTr", "WSTk")], na.rm = TRUE) * 1.05

# set up empty plot
plot(
  x = range_yr,
  y = c(0, ymax),
  type = "n",
  # xlab = "Year",
  xlab = NA,
  ylab = sprintf("Number reported (x %s)", div),
  xaxt = "n",
  yaxt = "n",
  las = 1
)

# get default tick labels for customization
# xtk <- axTicks(side = 1)
xtk <- seq(from = range_yr[1], to = range_yr[2], by = 2)
ytk <- axTicks(side = 2)

# custom x axis
axis(
  side = 1,
  at = xtk[xtk %% 2 == 1],
  labels = xtk[xtk %% 2 == 1],
  tck = -0.02#,
  # padj = -0.5
)

axis(side = 1, at = years, labels = NA, tck = -0.01)

# custom y axix
axis(side = 2, at = ytk, labels = ytk / 1000, las = 1)

# plot released WST
points(
  x = card_summary[["Year"]] - 0.08,
  y = card_summary[["WSTr"]],
  type = "h", lwd = 20, lend = 2,
  # col = "#00688B"
  col = "#4876FF"
)

# plot kept WST
points(
  x = card_summary[["Year"]] + 0.08,
  y = card_summary[["WSTk"]],
  type = "h", lwd = 20, lend = 2,
  # col = "#CD6600"
  col = "#4D4D4D"
)

mtext(
  text = "Source: Fishing Report Cards",
  side = 3,
  line = 0,
  adj = 0,
  col = "grey40"
)

```

# Report Card - White Sturgeon (released) 

```{r plot-rel_length}

bool_wstr <- with(data = AnglerCatch, expr = {
  Species %in% "White" & Fate %in% "released"
})

range_fl <- range(AnglerCatch[bool_wstr, "FL_cm"], na.rm = TRUE)

year_tick <- seq(from = range_yr[1], to = range_yr[2], by = 2)

plot(
  x = range_yr,
  y = range_fl,
  type = "n",
  xaxt = "n",
  yaxt = "n",
  # xlab = "Year",
  xlab = NA,
  ylab = "Fork length (cm)",
  las = 1
)

ytk <- axTicks(side = 2)

pt_col <- adjustcolor(
  col = gray(seq(from = 0,to = 0.5, length.out = 25)),
  alpha.f = 0.25
)

points(
  FL_cm ~ jitter(Year, factor = 1),
  data = AnglerCatch,
  pch = 1,
  cex = 1,
  subset = bool_wstr,
  col = pt_col
)

# stat data
bool_fate <- card_sp_length[["fate"]] %in% "released"

# se <- sqrt(wst_fl$FL[,"Var"] / wst_fl$FL[,"N"])
stdev <- sqrt(card_sp_length[bool_fate, "FL_cm.Var"])
lwr <- card_sp_length[bool_fate, "FL_cm.Avg"] - stdev
upr <- card_sp_length[bool_fate, "FL_cm.Avg"] + stdev
yr <- card_sp_length[bool_fate, "Year"]

segments(x0 = yr, y0 = lwr, y1 = upr, lwd = 2, col = "#CD3700")

points(
  FL_cm.Avg ~ Year,
  data = card_sp_length,
  pch = 19,
  cex = 2,
  subset = fate %in% "released",
  col = "#CD3700"
)

axis(side = 1, at = year_tick, labels = year_tick, tck = -0.02)
axis(side = 1, at = range_yr[1]:range_yr[2], labels = NA, tck = -0.01)
axis(side = 2, at = ytk, labels = ytk, tck = -0.01, hadj = 0.75, las = 1)

mtext(
  text = "Source: Fishing Report Cards",
  side = 3,
  line = 0,
  adj = 0,
  col = "grey40"
)

# points(
#   FL_cm.Med ~ Year,
#   data = card_sp_length,
#   pch = 0,
#   cex = 2,
#   subset = fate %in% "released",
#   col = "blue"
# )


```

# CPFV CPUE

```{r}

cpfv_year <- range(SFESuccessful[["Year"]])
cpfv_cpue <- range(SFESuccessful[["CPUE"]])


plot(
  x = cpfv_year,
  y = cpfv_cpue,
  type = "n",
  xaxt = "n",
  # yaxt = "n",
  # xlab = "Year",
  xlab = NA,
  ylab = "Catch per 100 angler-hour",
  las = 1
)

mtext(
  text = "Source: CPFV fishing logs",
  side = 3,
  line = 0,
  adj = 0,
  col = "grey40"
)

lines(
  CPUE ~ Year,
  data = SFESuccessful,
  lwd = 0.75,
  lty = 2,
  col = "grey40"
)
points(
  CPUE ~ Year,
  data = SFESuccessful,
  pch = 19,
  cex = 1.5,
  col = "royalblue3"
)

axis(
  side = 1,
  at = (cpfv_year[1]:cpfv_year[2])[cpfv_year[1]:cpfv_year[2] %% 2 == 0],
  labels = (cpfv_year[1]:cpfv_year[2])[cpfv_year[1]:cpfv_year[2] %% 2 == 0],
  tck = -0.02
)
axis(side = 1, at = cpfv_year[1]:cpfv_year[2], labels = NA, tck = -0.01)


```


# Mark-Recapture Length Distribution

```{r box-plot_FL, eval=FALSE}

Sturgeon <- readRDS("data/tagging/Sturgeon.rds")

year <- factor(Sturgeon$RelYear, levels = 1968:2017)

Sturgeon$Year <- factor(Sturgeon$RelYear, levels = 1954:2017)

test <- boxplot(FL ~ year, data = Sturgeon, subset = RelYear > 1967, drop = F)

plot(FL ~ RelYear, data = Sturgeon)

test$names <- as.numeric(test$names)
test$names <- 1968:2017

bxp(test)
abline(h = c(102, 152), col = 2)

boxplot(Sturgeon[, c("RelYear", "FL")])

```

```{r mr-len_freq, warning=FALSE}


wst_fl <- aggregate(
  formula = FL ~ RelYear,
  data = Sturgeon,
  # FUN = DescStat,
  FUN = function(x) unlist(DescStat(x)),
  # subset = Species %in% "White" &
  #   RelYear > 1967,
  na.action = na.pass
)

# se <- sqrt(wst_fl$FL[,"Var"] / wst_fl$FL[,"N"])
stdev <- sqrt(wst_fl$FL[,"Var"])
lwr <- wst_fl$FL[, "Avg"] - stdev
upr <- wst_fl$FL[, "Avg"] + stdev
yr <- wst_fl[["RelYear"]]

plot(
  x = range(yr),
  y = c(min(lwr), max(upr, 155)),
  type = "n",
  xaxt = "n",
  yaxt = "n",
  # xlab = "Year",
  xlab = NA,
  ylab = "Fork length (cm)",
  las = 1
)

ytk <- axTicks(side = 2)

mtext(
  text = "Source: CDFW",
  side = 3,
  line = 0,
  adj = 0,
  col = "grey40"
)

rect(
  xleft = min(yr) - 1,
  ybottom = 102,
  xright = max(yr) + 1,
  ytop = 152 ,
  col = "grey85",
  border = NA,
  xpd = NA
)
segments(x0 = yr, y0 = lwr, y1 = upr, col = "#27408B", lwd = 1.5)

points(
  x = yr,
  y = wst_fl[["FL"]][, "Avg"],
  cex = 1.5,
  col = "white",
  pch = 21,
  bg = "#27408B"
)

axis(
  side = 1,
  at = seq(from = 1970, to = 2017, by = 5),
  labels = seq(from = 1970, to = 2017, by = 5),
  tck = -0.02
)
axis(side = 1, at = 1968:2017, labels = NA, tck = -0.01)
axis(side = 2, at = ytk, labels = ytk, tck = -0.01, hadj = 0.75, las = 1)

# arrows()


```

# Mark-Recapture CPUE

```{r mr-cpue}

bool_yr <- CpueWST[["year"]] > 1967


cpue_ci <- with(data = CpueWST[bool_yr, ], expr = {
  
  se <- sqrt(v.cpue.slt / n)
  
  a <- 0.05 # alpha
  z <- qnorm(1 - a / 2)
  
  ci <- se * z
  
  l <- m.cpue.slt - ci
  u <- m.cpue.slt + ci
  
  list(
    lwr = l,
    upr = u,
    se = se,
    yrng = c(min(l, na.rm = TRUE), max(u, na.rm = TRUE))
  )
})

# yr <- wst_fl[["RelYear"]]
plot(
  x = range(CpueWST[["year"]]),
  y = cpue_ci[["yrng"]],
  type = "n",
  xaxt = "n",
  # xlab = "Year",
  xlab = NA,
  ylab = "Catch per 100 net-fathom hour",
  las = 1
)

mtext(
  text = "Source: CDFW",
  side = 3,
  line = 0,
  adj = 0,
  col = "grey40"
)
segments(
  x0 = CpueWST[bool_yr, "year"],
  y0 = cpue_ci[["lwr"]],
  y1 = cpue_ci[["upr"]],
  lwd = 5,
  col = "#27408B",
  lend = 2
)

axis(
  side = 1,
  at = seq(from = 1970, to = 2017, by = 5),
  labels = seq(from = 1970, to = 2017, by = 5),
  tck = -0.02
)
axis(side = 1, at = 1968:2017, labels = NA, tck = -0.01)
# points(
#   formula = m.cpue.slt ~ year,
#   data = CpueWST,
#   pch = 1,
#   cex = 0.5,
#   col = "royalblue3"
# )
# grid(nx = 10)

```

# Tag Returns (White Sturgeon)

```{r wst-harvest}

# for calculating return rate (# returned / # released)
harvest <- aggregate(
  formula = RetYear ~ RelYear + TagVal,
  data = WSTHarvest,
  FUN = function(x) {
    rl <- length(x)
    rt <- sum(x %in% 1)
    rate <- rt / rl
    c(Rel = rl, Ret = rt, Rate = rate)
  },
  na.action = na.pass
)

# for ease of further analytics
harvest <- data.frame(
  harvest[c("RelYear", "TagVal")],
  harvest[["RetYear"]]
)

# for plotting mean rate per tag value
mean_rate_val <- aggregate(
  formula = Rate ~ TagVal,
  data = harvest,
  FUN = mean
)

```

```{r return-rate}

# 
plot(
  formula = Rate ~ jitter(TagVal, factor = 0.5),
  data = harvest,
  col = "grey50",
  cex = 1.5,
  las = 1,
  ylab = "Rate",
  xlab = "Tag value ($)",
  yaxt = "n"
)

ytk <- axTicks(side = 2)

points(
  x = mean_rate_val[["TagVal"]],
  y = mean_rate_val[["Rate"]],
  col = "orange3",
  cex = 1.5,
  pch = 19
)

axis(
  side = 2,
  at = ytk,
  labels = format(ytk, digits = 2),
  hadj = 0.85,
  las = 1
)

mtext(
  text = "Source: CDFW",
  side = 3,
  line = 0,
  adj = 0,
  col = "grey40"
)

```

# Annual Return Rate

```{r return-rate2}

lkp_col_val <- c(
  `20` = "steelblue4",
  `50` = "tan2",
  `100` = "orange4",
  `150` = "red2"
)

lkp_pch_val <- c(
  `20` = "TW",
  `50` = "FF",
  `100` = "HH",
  `150` = "HF"
)

plot(
  formula = Rate ~ RelYear,
  data = harvest,
  pch = NA,
  las = 1,
  ylab = "Rate",
  # xlab = "Year",
  xlab = NA,
  xaxt = "n",
  yaxt = "n"
)

# xlbls <- axTicks(side = 1)
xlbls <- range(harvest[["RelYear"]])
xlbls <- seq(from = xlbls[1], to = xlbls[2], by = 2)

text(
  formula = Rate ~ RelYear,
  data = harvest,
  label = lkp_pch_val[as.character(harvest[["TagVal"]])],
  col = lkp_col_val[as.character(harvest[["TagVal"]])],
  cex = 1.0,
  family = "mono",
  font = 2
)

axis(side = 1, at = xlbls, labels = xlbls, tck = -0.02)
axis(side = 1, at = 1998:2017, labels = NA, tck = -0.01)

axis(
  side = 2,
  at = ytk,
  labels = format(ytk, digits = 2),
  hadj = 0.85,
  las = 1
)

mtext(
  text = "Source: CDFW",
  side = 3,
  line = 0,
  adj = 0,
  col = "grey40"
)

```

# Exploitation

```{r}

harvest_yr <- aggregate(
  formula = RetYear ~ RelYear,
  data = WSTHarvest,
  FUN = function(x) {
    rl <- length(x)
    rt <- sum(x %in% 1)
    rate <- rt / rl
    c(Rel = rl, Ret = rt, Rate = rate)
  },
  subset = !(TagVal  %in% 150),
  na.action = na.pass
)

# for ease of further analytics
harvest_yr <- data.frame(
  harvest_yr["RelYear"],
  harvest_yr[["RetYear"]]
)


exploitation <- Exploitation(
  rr = harvest_yr[["Ret"]],
  rt = harvest_yr[["Rel"]],
  repRate = 0.4,
  # tagL = 0.01,
  # tagM = 0.01
  tagL = 0,
  tagM = 0
)

```

```{r}


plot(
  x = range(harvest_yr[["RelYear"]]),
  y = range(exploitation[["Mu"]], exploitation[["MuPrime"]]),
  type = "n",
  las = 1, 
  # xlab = "Year",
  xlab = NA,
  ylab = "Rate",
  xaxt = "n",
  yaxt = "n"
)

ytk <- axTicks(side = 2)

# grid()

points(
  x = harvest_yr[["RelYear"]],
  y = exploitation[["Mu"]],
  col = "orange4",
  pch = 19,
  cex = 1.5
)

points(
  x = harvest_yr[["RelYear"]],
  y = exploitation[["MuPrime"]],
  col = "steelblue",
  pch = 19,
  cex = 1.5
)

points(
  x = 2015:2017,
  y = harvest[harvest$TagVal %in% 150, "Rate"],
  pch = "$"
)

mtext(
  text = sprintf("Response rate: %.2f", exploitation[["RepRate"]]),
  side = 3, adj = 1, col = "grey40"
)
mtext(
  text = "Source: CDFW",
  side = 3,
  line = 0,
  adj = 0,
  col = "grey40"
)


axis(side = 1, at = xlbls, labels = xlbls, tck = -0.02)
axis(side = 1, at = 1998:2017, labels = NA, tck = -0.01)
axis(
  side = 2,
  at = ytk,
  labels = format(ytk, digits = 2),
  hadj = 0.85,
  las = 1
)


```

# Model

```{r plot-lambda}

# x & y values for drawing polygon as lower & upper bounds
poly_list <- list(
  x = c(
    lambda_mu[["MuLevel"]][1],
    lambda_mu[["MuLevel"]],
    rev(lambda_mu[["MuLevel"]][-1])
  ),
  y = c(
    lambda_mu[["LBLambda"]][1],
    lambda_mu[["UBLambda"]],
    rev(lambda_mu[["LBLambda"]][-1])
  )
)

# create the plot with appropriate limits
plot(
  x = range(lambda_mu[["MuLevel"]]),
  y = range(lambda_mu[, c("LBLambda", "UBLambda")]),
  type = "n",
  panel.last = abline(h = 1, col = "grey50", lty = 2, lwd = 0.25),
  panel.first = polygon(poly_list, col = "grey90", border = NA),
  las = 1,
  xlab = "Exploitation",
  ylab = "Lambda",
  yaxt = "n"
)

lmd <- axTicks(side = 2)

# add data (mean lambda over mu)
lines(
  x = lambda_mu[["MuLevel"]],
  y = lambda_mu[["MeanLambda"]],
  col = "steelblue",
  lty = 1,
  lwd = 3
)

# optional: add current mu
# adding point might be better but would need to workout accurate y-val
# points(x = 0.13, y = 0.98, col = "darkorange", pch = 19)
abline(v = 0.13, col = "darkorange", lty = 2, lwd = 1)

axis(
  side = 2,
  at = lmd,
  labels = format(lmd, digits = 2),
  hadj = 0.85,
  las = 1
)

mtext(
  text = "Reference: Blackburn et al. 2018",
  side = 3,
  line = 0,
  adj = 0,
  col = "grey40"
)

# optional: lower & upper bounds as lines
# lines(
#   x = c(lambda_mu[["MuLevel"]]),
#   y = c(lambda_mu[["LBLambda"]]),
#   col = "blue"
# )
# lines(
#   x = c(lambda_mu[["MuLevel"]]),
#   y = c(lambda_mu[["UBLambda"]]),
#   col = "red"
# )

```


