---
title: "White Sturgeon Age-15 Abundance Estimates"
author: "California Department of Fish and Wildlife"
date: "April 19, 2016"
output: html_document
---

```{r File Paths and Libraries, echo=FALSE, warning=FALSE}

# file paths for absolute reference to source code and data connection
data_dir <- "C:/Data/jdubois/RDataConnections/Sturgeon/"

# load libraries as needed
library(ggplot2)
library(knitr)
# library(htmlwidgets)
# library(DT)
#library(reshape2)
#library(scales)

```

```{r Data, echo=FALSE, warning=FALSE, results='hide', message=FALSE, cache=TRUE}

source(file = paste0(data_dir, "OtherData.R"), echo = TRUE)
source(file = paste0(data_dir, "TaggingData.R"), echo = TRUE)

# load .rds abundance data (NOTE: .rds file is saved in Analysis-AltAbundance,
# so run code therein to refresh this .rds file - last refreshed 15-Apr-2016)
AltAbundance <- readRDS("AltAbund.rds")

```

```{r Source Code, echo=FALSE, warning=FALSE, message=FALSE}

# the source directory
source_dir <- "C:/Data/jdubois/RSourcedCode"

# sources needed functions and methods
source(file = paste(source_dir, "methods_len_freq.R", sep = "/"))
source(file = paste(source_dir, "methods_age_freq.R", sep = "/"))
source(file = paste(source_dir, "functions_global_data_subset.R", sep = "/"))
source(file = "source_stu_abund_age.R")

# some cleanup
rm(data_dir, source_dir)

```

```{r Variables, echo=FALSE}

# establish l-f breaks for WSTALKEY; note this alk uses TL not FL
#wst_alkey_breaks <- c(seq(from = 21, to = 186, by = 5), Inf)
wst_alkey_breaks <- c(wst_alkey$Bins, Inf)

```

### Background

The abundance of age-15 White Sturgeon is a key metric in the CPVIA's doubling goal. Historically, we (California Department of Fish and Wildlife, CDFW) have estimated age-15 abundance using mark-recapture data and the fraction of age-15 fish as calculated using an extant age-length key.

Herein we present age-15 White Sturgeon abundance using:

* Estimates of abundance using harvest and harvest rate  
* More recent age-length data (from US Fish and Wildlife Service, USFWS)  

### Methods & Results

We obtained length frequency distributions for White Sturgeon ≥ 85 cm TL from 2007-2014 (Fig. 1). (Minimum length of 85 cm TL set per investigations cited [here](https://nrm.dfg.ca.gov/FileHandler.ashx?DocumentId=114809).) Using age-length keys from CDFW (extant key) and one produced from recently-collected (~2014) USFWS age-length data, we obtained age frequency distributions for the same period (Fig. 2).

We estimated overall abundance (N) for White Sturgeon by summing the sub-legal (sub), legal (leg), and over-legal (ovr) abundance estimates as calculated per [CDFW unpublished](../Analysis-AltAbundance/StuAltAbundance.html). Using the fraction of age-15 White Sturgeon (AF~15~, as provided by the age frequency distributions), we calculated the abundance of age-15 fish (N~15~) using the equation below.  

${N}_{15} = {N}*{AF}_{15}$  

### Figures & Tabular Data

```{r LF Dist, echo=FALSE}

# NOTE: subsetting on >= 85 cm TL for use in getting at age-15 abundance. Alt
# abundance currently calculated using minimum WST length of >= 85 cm TL

# get length frequency distribution (from 2007-2014 in harmony with available 
# alt abundance estimates as of 19-Apr-2016); including only WST >= 85 cm TL & 
# *not* including recaptured WST (for now, this includes fish recorded as 
# shedding tag)
wst_lf_85 <- GetLenFreq(
  dat = SturgeonAll,
  colX = TL,
  by = "RelYear",
  breaks = wst_alkey_breaks,
  #seqBy = 5,
  intBins = TRUE,
  subset = Species %in% "White" &
    RelYear %in% 2007:2014 & TL >= 85 &
    # below removes recaptured WST
    StuType %in% c("Tag", "NoTag")
)

# to compare abundance at age-15 when calculated using >= 85 cm TL sized fish or
# when using (as in this section) fish within the current slot limit

# get lf dist for fish within slot (117-168 cm TL)
wst_lf_slot <- GetLenFreq(
  dat = SturgeonAll,
  colX = TL,
  by = "RelYear",
  breaks = wst_alkey_breaks,
  #seqBy = 5,
  intBins = TRUE,
  subset = Species %in% "White" &
    RelYear %in% 2007:2014 & TL %in% 117:168 &
    # below removes recaptured WST
    StuType %in% c("Tag", "NoTag")
)

```

```{r LF Plot, echo=FALSE, warning=FALSE, fig.height=8, fig.width=6, fig.cap="Fig. 1 White Sturgeon length frequency distributions 2007-2014. Length data from CDFW mark-recapture study. Blue vertical bar indicates median length. Length bins by 5 cm."}

# converting to factor for order when plotting
wst_lf_85$Data$LenCat <- factor(
  wst_lf_85$Data$LenCat,
  levels = c("sub", "leg", "ovr")
)

# plots length frequency distribution 2007-2014
plot(
  lf = wst_lf_85,
  lens = TL,
  fillBar = LenCat,
  type = "POT",
  addMed = TRUE,
  addN = TRUE
) + 
  facet_grid(facets = RelYear ~ .) +
  coord_cartesian(xlim = c(81, 186)) +
  scale_fill_manual(
    name = "",
    values = c(sub = "grey50", leg= "black", ovr = "steelblue")
  ) +
  xlab("Length bins (cm TL)") +
  theme_stu_lf

```

<br><br>

```{r Make ALKey, echo=FALSE}

# in this section we create an age-length key from USFWS data
wst_alk_usfws <- MakeALKey(
  dat = wst_usfws_age_len,
  len = TotalLength,
  age = Age,
  lenBreaks = wst_lf_85$Breaks,
  breakLabs = wst_lf_85$Bins,
  dia = FALSE
)

# convert wst_alk_usfws bins to integer for analytics to follow
wst_alk_usfws$Bins <- as.integer(as.character(wst_alk_usfws$Bins))

# confirm (if necessary)
# typeof(wst_alk_usfws$Bins)

```

```{r AF Dist, echo=FALSE}

# using different al keys get age frequency distribution for WST; for this
# purpose we get proportions of total (by year) for each age rather than count

# age frequency using extant CDFW age-length key
wst_af_cdfw <- GetAgeFreq(
  lfTable = wst_lf_85$Freq,
  alk = wst_alkey,
  prop = TRUE
)

# reshape for rbind()ing below (for now it seemed that reshaping was the best &
# quickest solutions for creating plot comparing af dist from these 2 al keys)
wst_af_cdfw_melt <- reshape2::melt(
  wst_af_cdfw,
  id.vars = "RelYear",
  variable.name = "Age",
  value.name = "Freq"
)

# adding source column for identity
wst_af_cdfw_melt$Source <- "CDFW"

# age frequency using 2014(?) USFWS age-length key
wst_af_usfws <- GetAgeFreq(
  lfTable = wst_lf_85$Freq,
  alk = wst_alk_usfws,
  prop = TRUE
)

# reshape for rbind()ing below
wst_af_usfws_melt <- reshape2::melt(
  wst_af_usfws,
  id.vars = "RelYear",
  variable.name = "Age",
  value.name = "Freq"
)

# adding source column for identity
wst_af_usfws_melt$Source <- "USFWS"

# age frequency using extant CDFW age-length key
wst_af_cdfw_slot <- GetAgeFreq(
  lfTable = wst_lf_slot$Freq,
  alk = wst_alkey,
  prop = TRUE
)

# reshape for rbind()ing below (for now it seemed that reshaping was the best &
# quickest solutions for creating plot comparing af dist from these 2 al keys)
wst_af_cdfw_slot_melt <- reshape2::melt(
  wst_af_cdfw_slot,
  id.vars = "RelYear",
  variable.name = "Age",
  value.name = "Freq"
)

# adding source column for identity
wst_af_cdfw_slot_melt$Source <- "CDFW"

# age frequency using data collected by USFWS
wst_af_usfws_slot <- GetAgeFreq(
  lfTable = wst_lf_slot$Freq,
  alk = wst_alk_usfws,
  prop = TRUE
)

# reshape for rbind()ing below
wst_af_usfws_slot_melt <- reshape2::melt(
  wst_af_usfws_slot,
  id.vars = "RelYear",
  variable.name = "Age",
  value.name = "Freq"
)

# adding source column for identity
wst_af_usfws_slot_melt$Source <- "USFWS"

# combine for plotting
wst_af <- rbind(wst_af_cdfw_melt, wst_af_usfws_melt)
wst_af_slot <- rbind(wst_af_cdfw_slot_melt, wst_af_usfws_slot_melt)

# clean up
rm(
  wst_af_cdfw_melt,
  wst_af_usfws_melt,
  wst_af_cdfw_slot_melt,
  wst_af_usfws_slot_melt
)

```

```{r AF Plot, echo=FALSE, fig.height=6, fig.width=8, fig.cap="Fig 2. White Sturgeon age frequency distributions 2007-2014 using two separate age-length keys - for fish >= 85 cm TL"}

# head(wst_af)

# converting for ease of displaying x-axis ticks
wst_af$Age <- as.numeric(as.character(wst_af$Age))

# ifelse(test = wst_af$Age %in% 15, yes = "red", no = "black")

# display age frequency distribution
ggplot(data = wst_af, mapping = aes(x = Age, y = Freq)) +
  # geom_bar(stat = "identity", fill = "black") +
  geom_bar(
    stat = "identity",
    fill = "black"
  ) +
  facet_grid(RelYear ~ Source) +
  scale_x_continuous(
    minor_breaks = seq(from = 0, to = 23, by = 1),
    expand = c(0.01, 0)
  ) +
  ylab("Fraction of total") +
  theme_stu_lf

```

<br>

```{r AF Plot2, echo=FALSE, fig.height=6, fig.width=8, fig.cap="Fig 3. White Sturgeon age frequency distributions 2007-2014 using two separate age-length keys - for slot sized fish only"}

# converting for ease of displaying x-axis ticks
wst_af_slot$Age <- as.numeric(as.character(wst_af_slot$Age))

# display age frequency distribution
ggplot( data = wst_af_slot, mapping = aes(x = Age, y = Freq)) +
  # geom_bar(stat = "identity", fill = "black") +
  geom_bar(
    stat = "identity",
    fill = "black"
  ) +
  facet_grid(RelYear ~ Source) +
  scale_x_continuous(
    minor_breaks = seq(from = 0, to = 23, by = 1),
    expand = c(0.01, 0)
  ) +
  ylab("Fraction of total") +
  theme_stu_lf

```

```{r Abund Age, echo=FALSE}

# AltAbundance

# get overall N (sub + leg + ovr)
annual_abun <- aggregate(
  formula = N ~ Year,
  data = AltAbundance,
  FUN = sum
)

# abundance for age 15 using cdfw af
abund_15_cdfw <- GetAbunAtAge(
  abun = annual_abun,
  af = wst_af_cdfw,
  age = 15
)

abund_15_cdfw$Source <- "CDFW"
abund_15_cdfw$Cat <- "all"

# abundance for age 15 using usfws af
abund_15_usfws <- GetAbunAtAge(
  abun = annual_abun,
  af = wst_af_usfws,
  age = 15
)

abund_15_usfws$Source <- "USFWS"
abund_15_usfws$Cat <- "all"

# age-15 abundance using slot-sized fish & CDFW age-length key
abund_15_slot_cdfw <- GetAbunAtAge(
  abun = AltAbundance[AltAbundance$LenCat %in% "leg", c("Year", "N")],
  af = wst_af_cdfw_slot,
  age = 15
)

abund_15_slot_cdfw$Source <- "CDFW"
abund_15_slot_cdfw$Cat <- "slot"

# age-15 abundance using slot-sized fish & USFWS age-length key
abund_15_slot_usfws <- GetAbunAtAge(
  abun = AltAbundance[AltAbundance$LenCat %in% "leg", c("Year", "N")],
  af = wst_af_usfws_slot,
  age = 15
)
abund_15_slot_usfws$Source <- "USFWS"
abund_15_slot_usfws$Cat <- "slot"

# combine for ease of plotting
abund_15 <- rbind(
  abund_15_cdfw,
  abund_15_usfws,
  abund_15_slot_cdfw[!is.na(abund_15_slot_cdfw$n_Age15), ],
  abund_15_slot_usfws[!is.na(abund_15_slot_usfws$n_Age15), ]
)

```

<br><br>

<caption>
  Tabular display of White Sturgeon age-15 abundance (CDFW - top 2, USFWS - bottom 2). Fraction data from age-length frequency distributions. (all = using fish ≥ 85 cm TL; slot = using fish 117-168 cm TL)
</caption>

```{r Tab Abund, echo=FALSE}

col_name_display <- c(
  "Year", "N (Total)",
  "Fraction", "N (Age 15)",
  "Length Category"
)
# capt <- "Age 15 abundance using %s age-length data"

# display tabular age-15 abundance
kable(
  abund_15_cdfw[ , -5],
  format = "markdown",
  row.names = FALSE,
  # padding = 5,
  col.names = col_name_display#,
  # caption = sprintf(capt, "CDFW")#,
  # format.args = list(big.mark = ',')
)

kable(
  abund_15_slot_cdfw[!is.na(abund_15_slot_cdfw$n_Age15), -5],
  format = "markdown",
  row.names = FALSE,
  # padding = 5,
  col.names = col_name_display#,
  # caption = sprintf(capt, "CDFW")#,
  # format.args = list(big.mark = ',')
)

kable(
  abund_15_usfws[ , -5],
  format = "markdown",
  row.names = FALSE,
  col.names = col_name_display#,
  # caption = sprintf(capt, "USFWS")
)

kable(
  abund_15_slot_usfws[!is.na(abund_15_slot_usfws$n_Age15), -5],
  format = "markdown",
  row.names = FALSE,
  col.names = col_name_display#,
  # caption = sprintf(capt, "USFWS")
)

```

<br>

```{r Plot Abun, echo=FALSE, fig.height=6, fig.width=8, fig.cap="Fig 4. White Sturgeon age-15 abundance 2007-2014 calculated from two different age-length sources"}

ggplot(
  data = abund_15,
  mapping = aes(
    x = as.character(Year),
    y = n_Age15
  )
) +
  geom_bar(
    mapping = aes(fill = Cat),
    stat = "identity",
    position = "dodge"
  ) +
  scale_y_continuous(
    labels = function(x) x / 1000,
    expand = c(0.01, 0)
  ) +
  scale_x_discrete(
    expand = c(0.01, 0)
  ) + 
  scale_fill_manual(
    name = "",
    values = c(all = "black", slot = "grey50")
  ) +
  labs(
    x = "Year",
    y = "Abundance - Age 15 (x 1000)"
  ) +
  facet_grid(facets = . ~ Source) +
  theme_stu_lf

```

<a href="#top">back to top</a>  
***  
Report ran: `r Sys.time()`  
End of report