---
title: |
  |  White Sturgeon Abundance Estimates
  |  from Harvest and Harvest Rate
author: "CDFW"
date: ''
output:
  pdf_document: default
  html_document: default
  github_document: default
header-includes: \usepackage{float} \usepackage{caption} \captionsetup[figure]{labelformat
  = empty}
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.pos = 'H'
)

knitr::opts_knit$set(root.dir = "~/RProjects/SturgeonPopMetrics/")

```

```{r libraries}

library(ggplot2)
library(knitr)
# library(reshape2)

```

```{r data}

# load all .rds data from card & tagging directories

# list.files(path = "data/card", pattern = ".rds$")
# list.files(path = "data/tagging", pattern = ".rds$")

# sourcing functions here to load data - likely will not use these functions
# elsewhere in this file

source(file = "source/functions_load_files.R")

LoadRdsFiles(dirName = "data/card")
LoadRdsFiles(dirName = "data/tagging")

SlotChanges <- read.csv(
  file = "data/SlotChanges.csv",
  header = TRUE,
  stringsAsFactors = FALSE
)

# clean up - functions for loading not needed
rm(list = ls.str(mode = "function"))

```

```{r source-files}

# source(file = "source/source_stu_hr.R")
source(file = "source/source_stu_mark-recap.R")
source(file = "source/functions_global_general.R")
source(file = "source/functions_global_record_count.R")
source(file = "source/functions_global_data_subset.R")
# source(file = "source/functions_cc_survival_rate.R")
# source(file = "source/methods_len_freq.R")
# source(file = "source/methods_age_freq.R")
# source(file = "source/source_stu_cc.R")
source(file = "source/source_alt_abundance.R")

```

```{r combine-card}

# this section combines the ALDS and archived (2007-2011) data

card_alds <- subset(
  SturgeonAlds,
  select = -MonthF
)

# adds Document (or Card) number - needed for reconciliation with tag return
# data
card_alds <- merge(
  card_alds,
  RetCardsAlds[, c("LicenseReportID", "DocumentNumber")],
  by = "LicenseReportID"
)

# adding date field for rbind()ing with 07 data
card_alds$DateOfCapture <- as.POSIXct(
  paste(
    card_alds$ItemYear,
    card_alds$Month,
    card_alds$Day,
    sep = "-"
  ),
  format = "%Y-%m-%d"
)

get_alds_cols <- c(
  "DocumentNumber", "ItemYear", "DateOfCapture", "LocCode",
  "SturgeonType", "Fate", "Length", "RewardDisk", "TL_cm", "FL_cm"
)

# subsetting & ordering for rbind()ing below
# card_alds <- card_alds[, c(11, 3, 12, 6, 9, 10, 7, 8)]
card_alds <- card_alds[, get_alds_cols]

# card_arch <- subset(
#   Sturgeon0711#,
#   # select = -c(SturgeonID, CorR)
#   # select = -c(CorR)
# )

# alias to avoid changing code below - fields SturgeonID, CorR are not in
# Sturgeon0711 so subsetting is not needed (28-Apr-2017)
card_arch <- Sturgeon0711

# set column names equal for rbind()ing
colnames(card_arch) <- colnames(card_alds)

card_all <- rbind(
  card_arch,
  card_alds
)

# clean up
rm(card_arch, card_alds, get_alds_cols)

```

```{r stats}

# adding year field for convenience
Effort$RelYear <- as.numeric(format(Effort$RelDate, "%Y"))

# rescue locations from April 2011 which we don't want for this analysis
drop_locations <- c("Fremont Weir", "Tisdale Bypass")

# get start, mid, and end dates
dates <- ApplyFunToDf(
  Effort,
  !(Location %in% drop_locations) &
    RelYear > 2005,
  splitVars = "RelYear",
  FUN = TaggingDates
)

# get length category stats for White Sturgeon from 2006-present (added
# 05-Apr-2016 total length cutoff in keeping with IEP article "Estimating Annual
# Abundance of White Sturgeon 85-116 and ≥ 169 Centimeters Total Length")
lc_stats <- ApplyFunToDf(
  SturgeonAll,
  Species %in% "White" & RelYear > 2005 & TL >= 85,
  splitVars = "RelYear",
  FUN = GetLcStats
)

# clean up
rm(drop_locations)

```

```{r abundance, comment=''}

# calculate alternative abundance estimates for White Sturgeon from
# 2007-present; output is a list so use GetAltAbundance() for cleaner display of
# abundance estimates with CIs
alt_abundance <- CalcAltAbundance(
  datHarvest = GetWstCount(datCard = card_all, dates = dates),
  datLcStats = lc_stats
)

# for convenience - variable used within narrative
max_year <- max(as.numeric(alt_abundance$Harvest$RelYear))

# tabular display of annual stats (e.g., harvest rate, tag returns) along with
# abundance estimates with Wald-type CIs and log-normal based CIs
stats_abund <- GetAltAbundance(dat = alt_abundance)

```

### Background
Herein are estimates of White Sturgeon abundance calculated using harvest (from Sturgeon Report Card) and harvest rate (from mark-recapture data). For more information and details regarding development of this method, please click [Abundance](https://nrm.dfg.ca.gov/FileHandler.ashx?DocumentId=43532).

We estimate abundance directly for legal-sized White Sturgeon, which from 2007 to present is 117-168 cm TL (centimeters total length; though from 2013 to present the equivalent is measured in fork length: 102-152 cm FL). Using proportions of catch and ratios, we estimate abundance for both sub-legal and over-legal sized White Sturgeon.

Note: any previous estimates have been updated using most-current (i.e., from `r Sys.Date()`) Report Card and mark-recapture data.

### Assigning Length Category
We assigned White Sturgeon --- caught during our mark-recapture field work --- to a length category of either 'sub' (sub-legal), 'leg' (legal), or 'ovr' (over-legal) based on measured length (total or fork) and data in the table below. The slot limit was instituted in 1990, thus sub-legal fish are below the slot limit and over-legal fish are above the slot limit. For this purpose, we use data in the last two rows of the table below.

```{r slot}

# changing Inf to NA for table display
SlotChanges[1, 5] <- NA

# outpout slot table for info on how LenCat was set
kable(
  SlotChanges,
  format = "pandoc",
  align = 'c',
  row.names = FALSE,
  col.names = c(
    "From Year",
    "To Year",
    "Species",
    "Min Length (cm)",
    "Max Length (cm)",
    "Length Type"
  ),
  caption = "Table of legal-size ranges for sturgeon from
  1954 to present. Note: in 2013 it became illegal to take
  Green Sturgeon. Species W = White & G = Green"
)

# reseting to Inf
SlotChanges[1, 5] <- Inf

```

For this purpose, we did not model length frequency distributions (as in [Gingras and DuBois 2015](https://nrm.dfg.ca.gov/FileHandler.ashx?DocumentId=114809)). Proportions and ratios were derived from non-modeled (i.e., straight) catch (see table below, where 'N' denotes count and 'P' denotes proportion). "SubToLeg" and "OvrToLeg" are ratios derived from the appropriate proportions. Also, per methodology described in [Gingras and DuBois 2015](https://nrm.dfg.ca.gov/FileHandler.ashx?DocumentId=114809), we limited White Sturgeon length to $\geqslant$ 85 cm TL. For example, in 2007 393 White Sturgeon were categorized as "sub," with n=34 < 85 cm TL. This yields NSub = 359.

Notes:

1. NAll = NSub + NLeg + NOvr
2. NUnk = number unknown
3. Proportions (P) calculated as N(Sub or Leg or Ovr) / NAll

```{r proportions}

#\u2265 = ≥

# display count and proportions table
kable(
  lc_stats,
  format = "pandoc",
  col.names = c(
    "Year", "NSub", "NLeg", "NOvr",
    "NAll", "NUnk", "PSub", "PLeg",
    "POvr", "SubToLeg", "OvrToLeg"
  ),
  caption = paste0(
    "Count and proportion of White Stureon ", 
    "from mark-recapture data, ",
    "where length limited to >= 85 cm TL"
  )
)

```

### Harvest
Since 2007, anglers fishing for sturgeon in California waters are required to report catch on a Sturgeon Fishing Report Card (Card). Annually, the Card provides --- among other information --- numbers of White Sturgeon kept (or harvest, ${H}$). Per developed methodology, we count harvest for three periods (start, middle, end) based on beginning and ending dates of our field season (August-October). The rationale is to use harvest more in line with harvest rate (the estimates of which use first-year tag returns). Because we have found harvest does not vary much by period (see table below), for abundance estimates we used harvest from the 'start' period.

Notes:

1. Each period has a 'from' and 'to' date (denoting one year) within which harvest is counted.
2. Harvest for 2006 not over a complete year, as Card not instituted until 2007.
3. Harvest for `r max_year` is preliminary, as data from `r  max_year + 1` Card not available until early `r max_year + 2`.
4. Harvest data are in BDSturgeonReportCard database (2007-2011) & ALDS (on-line) database (2012-present)

```{r harvest}

# display harvest dataframe
kable(
  alt_abundance$Harvest,
  format = "pandoc",
  #align = 'c',
  row.names = FALSE,
  col.names = c(
    "Year",
    "Period",
    "From",
    "To",
    "Harvest"
  ),
  caption = "Harvest of White Sturgeon as reported on
  Cards from 2006-present. Harvest for 2006 included FIO
  and is partial, as Card not instituted until 2007.",
  format.args = list(big.mark = ',')
)

```

### Harvest Rate
We estimate harvest rate ($\hat{h}$) annually using angler tag returns (Ret or returns) and CDFW tag releases (Rel or releases). To minimize bias, we only use first-year tag returns (i.e., tagged fished caught within one year of release; denoted ${_f}$ in equation below). Below are harvest rate estimates for White Sturgeon legal-sized when tagged. The returns and releases presented in the table are for this demographic. Period, dates (from & to), and harvest included in table for convenience.

$\hat{h} = \frac{Ret_f}{Rel}$

Notes:

1. Rate for 2006 provided for information only.
2. Rate for `r max_year` is preliminary and will change with future data (returns).
3. Returns (Ret) includes tags physically returned to us by angler and tag reported on Cards but not physically returned (Ret & Rel data are in BDSturgeonTagging database)
4. We release reward tags (current values: $50; $100; $150) but do not consider reward here (i.e., for the harvest rate estimate). See [Gingras and DuBois 2014](https://nrm.dfg.ca.gov/FileHandler.ashx?DocumentId=74460) for further information regarding bias in annual harvest rate estimates. 

```{r harvest-rate}

# display harvest rate dataframe
kable(
  stats_abund$AnnualStats,
  format = "pandoc",
  digits = 3,
  #align = 'c',
  row.names = FALSE,
  col.names = c(
    "Year",
    "Period",
    "From",
    "To",
    "Harvest",
    "Ret",
    "Rel",
    "Rate"
  ),
  caption = "Harvest rate estimates of White Sturgeon
  legal-sized when tagged (2006-present). Note: rates
  for 2016 will change with further data.",
  format.args = list(big.mark = ',')
)

```

### Abundance Estimates

Below are (alternative) abundance estimates ($\hat{N}$) for White Sturgeon for the three length categories. Provided herewith are standard error (SE), Wald-type confidence intervals (CI), and log-normal confidence intervals (lower = LNLB, upper = LNUB). (Confidence interval algorithms provided courtesy of K. Newman & L. Mitchell of USFWS.)

**Abundance equations**  
$\hat{N}_{leg} = \frac{H}{\hat{h}}$  
$\hat{N}_{sub} = \hat{N}_{leg} * SubToLeg$  
$\hat{N}_{ovr} = \hat{N}_{leg} * OvrToLeg$  

**Standard error and variance**  
${SE}({\hat{N}_{leg}}) = \frac{H}{\sqrt{\hat{h}^{3} * {Rel}}}$  
${VAR}({\hat{N}_{leg}}) = {SE}({\hat{N}_{leg}})^2$  

${SE}({\hat{N}_{sub}}) = \sqrt{{VAR}({\hat{N}_{leg}}) * SubToLeg}$  
${SE}({\hat{N}_{ovr}}) = \sqrt{{VAR}({\hat{N}_{leg}}) * OvrToLeg}$  

**Wald-type confidence interval (CI)**  
${CI}(\hat{N}) = {\pm}{Z} * {SE}(\hat{N})$  

**Log-normal confidence intervals (lower & upper bounds)**  
${LNLB}(\hat{N}) = \hat{N} * exp({-Z} * {\frac{SE(\hat{N})}{\hat{N}}})$  
${LNUB}(\hat{N}) = \hat{N} * exp({Z} * {\frac{SE(\hat{N})}{\hat{N}}})$  

```{r abundance-tabular}

# for display purpose in table below
stats_abund$Abundance$Year <- as.character(stats_abund$Abundance$Year)

# display abundance estimates
kable(
  stats_abund$Abundance,
  format = "pandoc",
  digits = 0,
  #align = 'c',
  row.names = FALSE,
  col.names = c(
    "Year",
    "Category",
    "N",
    "SE",
    "CI",
    "LNLB",
    "LNUB"
  ),
  caption = "Annual White Sturgeon abundance estimates by size category",
  format.args = list(big.mark = ',')
)

# resetting numeric for plotting
stats_abund$Abundance$Year <- as.numeric(stats_abund$Abundance$Year)

```

\pagebreak

Plot of abundance estimates with log-normal confidence intervals. Note: estimates for `r max(stats_abund$Abundance$Year)` available ~`r max(stats_abund$Abundance$Year) + 2`.

```{r graphics, fig.height=7, fig.width=5, fig.align='center', fig.cap="White Sturgeon abundance estimates from harvest and harvest rate with log-normal confidence intervals; note: separate y-axis for each category"}

stats_abund$Abundance$LenCat <- factor(
  stats_abund$Abundance$LenCat,
  levels = c(sub = "sub", leg = "leg", ovr = "ovr")
)

year_range <- range(stats_abund$Abundance$Year)

# possilbe plotting option (change data source)
plt <- ggplot(
  data = stats_abund$Abundance[stats_abund$Abundance$Year < year_range[2], ],
  mapping = aes(x = as.character(Year), y = N)
)

plt +
  geom_bar(stat = "identity", fill = "black") +
  #facet_wrap(facets = ~LenCat, scales = "free_y", ncol = 1)
  facet_grid(facets = LenCat ~ ., scales = "free_y") +
  geom_errorbar(
    mapping = aes(ymax = LNUB, ymin = LNLB),
    colour = "steelblue",
    size = 0.5,
    width = 0.1
  ) +
  scale_y_continuous(
    name = "Abundance (x 1000)",
    #name = quote(Abundance * 10^3),
    labels = function(x) x / 1000,
    expand = c(0.02, 0)
  ) +
  xlab("Year")


  # scale_y_continuous(
  #   minor_breaks = 5,
  #   labels = function(x) x / 1000
  # ) +
  #ylab("Abundance (x 1000)")

```

***
Report ran: `r Sys.time()`  
End of report