---

---

```{r setup, include=FALSE}

knitr::opts_knit$set(root.dir = "~/RProjects/SturgeonPopMetrics")
knitr::opts_chunk$set(echo = FALSE)

```

```{r libraries, warning=FALSE, message=FALSE, echo=FALSE}

# source(file = "~/RSourcedCode/methods_desc_stats.R")

library(dplyr)

```

```{r load-data, echo=FALSE}

SturgeonAlds <- readRDS(file = "data/card/SturgeonAlds.rds")
Sturgeon0711 <- readRDS(file = "data/card/Sturgeon0711.rds")
RetCardsAlds <- readRDS(file = "data/card/RetCardsAlds.rds")
PurCardsAlds <- readRDS(file = "data/card/PurCardsAlds.rds")

count_alds <- nrow(SturgeonAlds)
count_0711 <- nrow(Sturgeon0711)

# summary(SturgeonAlds)
# summary(Sturgeon0711)

```

```{r somevars, echo=FALSE}

col_names <- c(
  "Id", "Year", "Month", "LocCode", "Fate",
  "Length", "TagNum", "FL_cm", "TL_cm"
)


```

```{r data-subset, echo=FALSE}

# subset data on white stu & prepare for rbind()ing

# usefule for seasonal summary
Sturgeon0711$Month <- as.numeric(format(Sturgeon0711[["DateOfCapture"]], "%m"))

arch_wst <- subset(
  Sturgeon0711,
  subset = Species %in% "White",
  select = c(
    CardNum, Year, Month, LocationCode,
    Fate, Length, TagNum, FL_cm, TL_cm
  )
)

colnames(arch_wst) <- col_names

alds_wst <- subset(
  SturgeonAlds,
  subset = SturgeonType %in% "White",
  select = c(
    CustomerID, ItemYear, Month, LocCode,
    Fate, Length, RewardDisk, FL_cm, TL_cm
  )
)

colnames(alds_wst) <- col_names


card <- rbind(arch_wst, alds_wst)
rm(arch_wst, alds_wst)

```

Summary data about number of White Sturgeon kept and released each year.

```{r}

catch <- aggregate(
  formula = Id ~ Year + Fate,
  data = card,
  FUN = length
)

catch$Fate <- factor(catch$Fate, levels = c("kept", "released"))
colnames(catch) <- c("Year", "Fate", "NumOfWst")

anglers <- aggregate(
  formula = Id ~ Year,
  data = card,
  FUN = function(x) length(unique(x))
)

colnames(anglers) <- c("Year", "NumOfAnglers")

as.data.frame(catch)
as.data.frame(anglers)

```

Plot of data above.

```{r echo=FALSE}

# plot(Id ~ Year, data = anglers)

par(mar = c(3.5, 4, 1, 4) + 0.1, bty = "u")

plot(
  x = anglers[["Year"]],
  y = anglers[["NumOfAnglers"]],
  type = "h",
  lwd = 10,
  lend = 2,
  col = "grey75",
  ylim = c(0, max(anglers[["NumOfAnglers"]])),
  ylab = "Number of anglers (x 1000)",
  xlab = NA,
  yaxt = "n",
  xaxt = "n"
)

# range(anglers[["Id"]])

axis(
  side = 2,
  at = seq(0, 3000, 500),
  labels = format(seq(0, 3000, 500) / 1000),
  las = 1
)

par(new = TRUE)

# catch <- catch[order(catch[["Year"]]), ]

# points(Id ~ Year, data = catch, type = "b", subset = Fate %in% "kept", ylim = c(0, 6000))
# points(Id ~ Year, data = catch, type = "b", subset = Fate %in% "released", ylim = c(0, 6000))

bool <- catch[["Fate"]] %in% "released"

plot(
  x = catch[["Year"]],
  y = catch[["NumOfWst"]],
  type = "n",
  # col = as.numeric(catch[["Fate"]]),
  # lty = as.numeric(catch[["Fate"]]),
  
  axes = FALSE,
  ylab = NA,
  xlab = NA
)

# range(catch[["Id"]])

pts <- lapply(c("kept", "released"), FUN = function(x) {
  
  color <- c(kept = "steelblue", released = "darkorange")
  
  points(
    formula = NumOfWst ~ Year,
    data = catch,
    subset = Fate %in% x,
    type = "b",
    col = color[x],
    cex = 1.1
  )
})


axis(
  side = 4,
  at = seq(1400, 6200, 700),
  labels = ( seq(1400, 6200, 700)) / 100,
  las = 2
  # lwd.ticks = ,
  # hadj =
)

mtext(text = "Catch (x 100)", side = 4, line = 2.5)
mtext(text = "Year", side = 1, line = 2)

axis(
  side = 1,
  at = (2007:2017)[(2007:2017) %% 2 == 1],
  labels = (2007:2017)[(2007:2017) %% 2 == 1],
  tck = -0.03,
  padj = -0.5
  # las = 2
  # lwd.ticks = 5
  # hadj =
)

axis(
  side = 1,
  at = 2007:2017,
  labels = NA,
  tck = -0.01
)

legend(#x = 2009,
       #y = 6300,
     "topleft",
    # fill = c("grey75", "steelblue"),
    col = c("grey75", "steelblue", "darkorange"),
    pch = c(15, 1, 1),
    legend = c("anglers", "kept", "released"),
    pt.cex = 1.5, ncol = 3, box.col = "grey90", border = NA)



```

Some angler demographics...these data only available from 2012 to present.

Count of cards issued (or purchased beginning 2013) by year and gender along with median age.

```{r angler-info, echo=FALSE}

angler_info <- PurCardsAlds %>% 
  group_by(ItemYear, Gender = factor(Gender, levels = c('M', 'F'))) %>% 
  filter(!duplicated(CustomerID) & ItemYear < 2018) %>% 
  summarise(
    Count = n(),
    MeadianAge = median(2018 - BirthYear)
  )
  
as.data.frame(angler_info)

```
 
Fraction of top 5 counties by year. County = where angler resides. Fractions don't vary much annually and majority of sturgeon anglers is local (to the Delta - Estuary).

```{r county-info, echo=FALSE}

# 
years <- setNames(object = 2012:2017, nm = 2012:2017)

county_info <- vapply(years, FUN = function(y) {
  
  d <- PurCardsAlds[PurCardsAlds[["ItemYear"]] %in% y, ]
  d <- d[!duplicated(d[["CustomerID"]]), ]
  
  county <- prop.table(table(d[["County"]], useNA = "ifany"))
  
  sort(county, decreasing = TRUE)[1:5]
  
}, FUN.VALUE = numeric(5L))

t(county_info)

```

More anglers now appear to be submitting on-line compared to when ALDS came on-line in 2012.

IS = Internet submission (by angler)  
CC = Control Center (by CDFW staff)

```{r alds-returns, echo=FALSE}

# head(RetCardsAlds)
# 
# table(RetCardsAlds[["Code"]], useNA = "ifany")
# table(RetCardsAlds[["CustomerSourceCode"]], useNA = "ifany")

ret_cards_alds <- RetCardsAlds %>% 
  filter(ItemYear < 2018) %>% 
  group_by(ItemYear, CustomerSourceCode) %>% 
  summarise(Count = n())

as.data.frame(ret_cards_alds)

```


