---

---

```{r setup, include=FALSE}

knitr::opts_knit$set(root.dir = "~/RProjects/SturgeonPopMetrics/")

```

## Introduction

Herein, we explain the complexities of estimating White Sturgeon harvest rate (exploitation, $\mu$), a componet of mortality. We calculate these estimates using tag releases (by CDFW) and angler tag returns (returns).

## Libraries

We load the `sportfish` package, currently available on GitHub. For now (27-Feb-2019), this is the only package required.

```{r load-libraries}

library(sportfish)
# library(package)

```

<!--
mortality
- fishing
- natural
- total

use Ricker eqn to get from one metric to another

for exploitation need tags released by reward, tags returned (first year)
-->

## Load Data

We load all `.rds` files from directory `data/tagging`. To keep our workspace clean, we load these files into a new environment called `Tagging`.

```{r load-data}

# the data directory for bay study
data_dir <- "data/tagging"

# list.files(path = data_dir, pattern = ".rds")

Tagging <- new.env()

ReadRDSFiles(fileDir = data_dir, envir = Tagging)

# for consumer price index (adjusting tag $ for inflation) from
# https://www.measuringworth.com/datasets/uscpi/result.php
cpi <- read.csv(file = "~/RProjects/1_DataFiles/USCPI_1998-2018.csv")
cpi$Ratio <- cpi[["USCPI"]][1] / cpi[["USCPI"]]
cpi$Ratio2 <- cpi[["USCPI"]][nrow(cpi)] / cpi[["USCPI"]]

# clean up
rm(data_dir)

```

For analytics herein, we'll use `WSTHarvest`, which has currently `r nrow(Tagging[["WSTHarvest"]])` records. We restrict our analytics to using only data from 1998 onward. In 1998, the CDFW began issuing three tag reward denominations: \$20; \$50; and \$100.

## Variables

Here we create some variables we'll use throughout this process. We create them here and now for convenience.

*Note*: none needed at this time.

```{r variables}


```

```{r eval=FALSE, include=FALSE}

# head(Tagging$WSTHarvest)

# count of "blank" disc tags
with(data = Tagging$WSTHarvest, expr = {
  r <- TagNum == ""
  # r <- TagNum == "" & !is.na(PitTag)
  # r <- TagNum == "" & is.na(PitTag)
  s <- split(r, f = RelYear)
  vapply(s, FUN = sum, FUN.VALUE = numeric(1L))
})

# count of tags released
with(data = Tagging$WSTHarvest, expr = {
  s <- split(TagNum, f = RelYear)
  vapply(s, FUN = length, FUN.VALUE = numeric(1L))
})

# count of tag values by year
with(data = Tagging$WSTHarvest, expr = {
  table(RelYear, TagVal, useNA = "ifany")
})

# count of tag values by year
with(data = Tagging$WSTHarvest, expr = {
  table(RelYear, RetYear, useNA = "ifany")
})

# 
with(data = Tagging$WSTHarvest, expr = {
  table(RetFate, useNA = "ifany")
})

range(Tagging$WSTHarvest$RelDate)

```

## Method

<Narrative here>

#### Return Rate

We need to get per year per tag value, released count, return count, and return rate. For this purpose, we'll use all data (post 1997) irrespective of size (i.e., fork length) and (or) length category (e.g., legal-sized when tagged). Return rate is simply number tags returned divided by number tags released.

```{r harvest}

# for calculating return rate (# returned / # released)
harvest <- aggregate(
  formula = RetYear ~ RelYear + TagVal,
  data = Tagging[["WSTHarvest"]],
  FUN = function(x) {
    rl <- length(x)
    rt <- sum(x %in% 1)
    rate <- rt / rl
    c(Rel = rl, Ret = rt, Rate = rate)
  },
  # subset = LenCat %in% "leg",
  # subset = LenCat %in% "leg" & RelYear > 1997,
  subset = RelYear > 1997,
  na.action = na.pass
)

# for ease of further analytics
harvest <- data.frame(
  harvest[c("RelYear", "TagVal")],
  harvest[["RetYear"]]
)

```

#### Return Rate by Tag Value

We can display return rate a couple of different ways. Here we display return rate as a function of tag value (adding some noise to minimize overplotting). We then overlay mean values (orange point).

```{r plot-return_rate}

# for plotting mean rate per tag value
mean_rate_val <- aggregate(
  formula = Rate ~ TagVal,
  data = harvest,
  FUN = mean
)

# 
plot(
  formula = Rate ~ jitter(TagVal, factor = 0.5),
  data = harvest,
  col = "grey50",
  cex = 1.5,
  las = 1,
  ylab = "Return rate",
  xlab = "Tag value"
)

points(
  x = mean_rate_val[["TagVal"]],
  y = mean_rate_val[["Rate"]],
  col = "orange3",
  cex = 1.5,
  pch = 19
)

```

Here we plot return rate (by dollar value) over time (release or tagging year). We label dollar values as 'TW' (\$20), 'FF' (\$50), 'HH' (\$100), and 'HF' (\$150).

```{r plot-return_rate2}

lkp_col_val <- c(
  `20` = "steelblue4",
  `50` = "tan2",
  `100` = "orange4",
  `150` = "red2"
)

lkp_pch_val <- c(
  `20` = "TW",
  `50` = "FF",
  `100` = "HH",
  `150` = "HF"
)

plot(
  formula = Rate ~ RelYear,
  data = harvest,
  pch = NA,
  las = 1,
  ylab = "Return rate",
  xlab = "Year"
)

# mtext(text = paste0(lkp_pch_val, collapse = " | "), side = 3, adj = 0)

text(
  formula = Rate ~ RelYear,
  data = harvest,
  label = lkp_pch_val[as.character(harvest[["TagVal"]])],
  col = lkp_col_val[as.character(harvest[["TagVal"]])],
  cex = 1.0,
  family = "mono",
  font = 2
)

```

#### Inflation Adjustment

We've plotted return rate by dollar value and over time to eventually get some measure of non-reporting. Below we do likewise, this time adjusting for inflation using the Consumer Price Index.

Work in progress

```{r eval=FALSE, include=FALSE}

# index matching based on year (release year in harvest) to use correct ratio
# adjuster in cpi
iyr <- match(x = harvest[["RelYear"]], table = cpi[["Year"]])

harvest[["TagVal"]]
harvest[["TagVal"]] * cpi[["Ratio"]][iyr]

plot(x = harvest[["TagVal"]] / cpi[["Ratio2"]][iyr], y = harvest[["Rate"]])
abline(h = 0.075, col = 2)

test <- vector(length = 48)
test[harvest[["Rate"]] > 0.075] <- TRUE
sum(test)

test <- data.frame(
  Binom = test,
  Val = harvest[["TagVal"]] / cpi[["Ratio2"]][iyr]
)

test <- data.frame(
  Dollar = c(20, 50, 100, 150),
  Rate = rep_rate
)

# g <- glm(Binom ~ Val, data = test, family = binomial(link = "logit"))
g <- glm(Rate ~ Dollar, data = test, family = binomial(link = "logit"))

# plogis(predict(object = g, newdata = list(20:140), type = "link"))
gprd <- predict(object = g, newdata = data.frame(Dollar=20:140), type = "response")

plot(test$Dollar, test$Rate)
lines(20:250, gprd, col = "blue")

coefficients(g)

x <- test$Val
exp(-6.58669790 + (0.06620434*x)) / (1 + exp(-6.58669790 + (0.06620434*x)))

```

#### Reporting Rate

Here we try a couple different approaches to calculate reporting rate. We'll use reporting rate to adjust for non-response.

We are still in the "tinkering" process, and for now we'll assume 100% reporting of the \$150 tag. As such, we can divide each mean return rate by the maximum mean return rate (in this case the \$150 tag). Then, we take the average (n=3; \$20, \$50, and \$100 tags).

```{r rep-rate1}

rep_rate <- mean_rate_val[["Rate"]] / max(mean_rate_val[["Rate"]])

mean_rep_rate <- mean(rep_rate[1:3])
mean_rep_rate

```

Our second approach assumes the \$150 tag as the "reward" tag and all others as "non-reward." There are some inherent problems with this, but typically mark-recapture studies issue a high-dollar reward mixed with non-reward tags. Then the reporting rate is simply the non-reward return rate divided by the reward return rate.

With this approach, we can use `sportfish::ReportingRate()` created per methodolody in (Idaho tag return study). We see both methods produce similary results.

```{r rep-rate2}

lkp_dol <- c(`20` = "NR", `50` = "NR", `100` = "NR", `150` = "RR")

# return release as non-reward reward
ret_rel_nrr <- aggregate(
  formula =cbind(Ret, Rel) ~ lkp_dol[as.character(TagVal)],
  data = harvest,
  FUN = sum
)

colnames(ret_rel_nrr)[1] <- "Val"
attr(ret_rel_nrr[["Val"]], which = "names") <- NULL

rep_rate_nrr <- with(data = ret_rel_nrr, expr = {
  bn <- Val %in% "NR"
  br <- Val %in% "RR"
  ReportingRate(rr = Ret[bn], rt = Rel[bn], nr = Ret[br], nt = Rel[br])
})

rep_rate_nrr[["RepRate"]]
mean_rep_rate

```

#### Annual Harvest Rate

Harvest rate can be biased low if we do not account for non-reporting. Simply using return rate will in most cases (i.e., when assuming 100% reporting of all tags) yield a false sense of low harvest.

Here we present "raw" harvest rate (i.e., no adjustment) and adjusted harvest rate using ~0.4 response rate. We use `sportfish::Exploitation()` developed per methodoloy in <Idaho reference paper>. We limit our data to only \$20, \$50, and \$100 tags, as we are using the \$150 to adjust for non-response.

```{r harvest-rate}

harvest_yr <- aggregate(
  formula = RetYear ~ RelYear,
  data = Tagging[["WSTHarvest"]],
  FUN = function(x) {
    rl <- length(x)
    rt <- sum(x %in% 1)
    rate <- rt / rl
    c(Rel = rl, Ret = rt, Rate = rate)
  },
  # subset = LenCat %in% "leg",
  # subset = LenCat %in% "leg" &
  #   RelYear > 1997 &
  #   !(TagVal  %in% 150),
  subset = RelYear > 1997 &
    !(TagVal  %in% 150),
  na.action = na.pass
)

# for ease of further analytics
harvest_yr <- data.frame(
  harvest_yr["RelYear"],
  harvest_yr[["RetYear"]]
)

```

We adjust for tag loss and tagging mortality using function parameters `tagL` and `tagM`. We set both to 0.01, but we should tinker with these inputs to observe effects on final result.

```{r exploitation}

exploitation <- Exploitation(
  rr = harvest_yr[["Ret"]],
  rt = harvest_yr[["Rel"]],
  repRate = round(
    max(mean_rep_rate, rep_rate_nrr[["RepRate"]]),
    digits = 2
  ),
  tagL = 0.01,
  tagM = 0.01
)

```

We plot annual harvest rate for both "raw" and adjust harvest rate estimates. The \$ represents the "raw" \$150 rate.

*Note*: The exceptionally high rate for 2016 can be explained somewhat be the very low number of tags released that year. 

```{r plot-exploitation}

plot(
  x = range(harvest_yr[["RelYear"]]),
  y = range(exploitation[["Mu"]], exploitation[["MuPrime"]]),
  type = "n",
  las = 1, 
  xlab = "Year",
  ylab = "Harvest rate"
)

grid()

points(
  x = harvest_yr[["RelYear"]],
  y = exploitation[["Mu"]],
  col = "orange4",
  pch = 19,
  cex = 1.5
)

points(
  x = harvest_yr[["RelYear"]],
  y = exploitation[["MuPrime"]],
  col = "steelblue",
  pch = 19,
  cex = 1.5
)

points(
  x = 2015:2017,
  y = harvest[harvest$TagVal %in% 150, "Rate"],
  pch = "$"
)

mtext(
  text = sprintf("RepRate: %.2f", exploitation[["RepRate"]]),
  side = 3, adj = 0
)

```

## Further Tinkering

We need some reliable measure of exploitation (i.e., rate at which fish are being removed from the fishery). Heretofore, we've inclduded all tagged White Sturgeon, including those that might be sub- or over-legal and those where the angler reported releasing said fish. Not every angler reports the fate (i.e., kept or released) of his/her catch.

#### Fork Length vs. Days at Large

One alternative might be to look at length at tagging versus days at large. This will perhaps give some sense of fish elligible for harvest (i.e., within legal limits) and might provide some measure of restricting our data based on size range. (Blackburn et al. 2018 accounted for growth into and out of the slot limit using available age-at-length growth data.) Below we display fork length as a function of days-at-large (i.e., days post tagging before catch) color coding data based on fate: kept = orange; blue = released; grey = not reported.

```{r dal-length}

dal <- TagRetYear(
  data = Tagging[["WSTHarvest"]],
  dcap = RetDate,
  drel = RelDate
)[["DAL"]]

lkp_col_fate <- c(
  kept = "orange3",
  relAlive = "royalblue4",
  unk = "grey50"
)

lkp_pch_fate <- c(
  kept = 16,
  relAlive = 15,
  unk = 1
)

plot_range <- with(data = Tagging[["WSTHarvest"]], expr = {
  b <- RelYear > 1997 & RetYear %in% 1
  
  list(
    X = range(dal[b], na.rm = TRUE),
    Y = range(RelFL[b], na.rm = TRUE)
  )
})

plot(
  x = plot_range[["X"]],
  y = plot_range[["Y"]],
  type = "n",
  las = 1,
  xlab = "Days at large",
  ylab = "Fork length (cm)"
)

# rect(
#   xleft = plot_range[["X"]][1] - 5,
#   ybottom = 102,
#   xright = plot_range[["X"]][2] + 5,
#   ytop = 152 ,
#   col = "grey85",
#   border = NA,
#   xpd = NA
# )

points(
  formula = RelFL ~ dal,
  data = Tagging[["WSTHarvest"]],
  subset = RelYear > 1997 & RetYear %in% 1,
  col = lkp_col_fate[Tagging[["WSTHarvest"]][["RetFate"]]],
  pch = lkp_pch_fate[Tagging[["WSTHarvest"]][["RetFate"]]],
  cex = 1.5
)

```

#### Fish Disposition (Fate)

Below we display by release (tagging year) the count of tag returns by fate (kept, released, or unknown). We could use only fish reported as "kept" when estimating exploitation. However, this likely will marked reduce sample size.

```{r tbl-year_fate}

with(data = Tagging[["WSTHarvest"]], expr = {
  b <- RetYear %in% 1 & RelYear > 1997
  table(RelYear[b], RetFate[b], useNA = "ifany")
})

```

## Modeling (Random Sampling)

Section in process. Attempting to gain some insight into distribution of returns & if it's possible to expand our dataset using sampling techniques. Or can we employ some predictive model to ultimately get at reporting rate and thus exploitation.

```{r rate-freq, eval=FALSE}

rate_freq <- Frequency(harvest[["Rate"]], binWidth = 0.01)
plot(rate_freq)
lines(rpois(50, lambda = seq(0, 0.5, by = .1)))

# pnorm()

ret_freq <- Frequency(harvest[["Ret"]], binWidth = 1)
plot(ret_freq)
lines(dpois(0:50, lambda = 5.6875), col = "steelblue")
lines(dpois(0:50, lambda = 3), col = "green")
lines(dnorm(0:20, mean = 5.6875, sd = sqrt(16.43218)))
# lines(density(rpois(50, lambda = 4.5)))


# curve(density(rpois(48, lambda = mean(x))))

ret_freq$xstats()
rate_freq$xstats()

```

---

CDFW, SportFish Unit  
`r Sys.Date()`  
