---

---

```{r setup, include=FALSE}

#knitr::opts_knit$set(root.dir = "~/RProjects/SturgeonPopMetrics/")
knitr::opts_knit$set(root.dir = "U:/SportFish/Staff Files/JDuBois/0_RProjects/SturgeonPopMetrics/")
now <- Sys.Date()

```

## Introduction

<!-- Herein, we calculate White Sturgeon absolute abundance using CDFW mark-recaptured data. Also, <GREEN STURGEON??> -->

## Libraries

We load the `sportfish` package, currently available on GitHub. For now (`r now`), this is the only package required.

```{r load-libraries}

library(sportfish)
# library(package)

```

## Load Data

We load all `.rds` files from directory `data/tagging`. To keep our workspace clean, we load these files into a new environment called `Tagging`.

```{r load-data}

# the data directory for bay study
data_dir <- "data/tagging"

# list.files(path = data_dir, pattern = ".rds")
Tagging <- new.env()

ReadRDSFiles(fileDir = data_dir, envir = Tagging)

notes <- readLines(file.path(data_dir, "data-log"), n = 2)
notes <- paste0(notes[1], notes[2])

# clean up
rm(data_dir)

```

**Note**: `r notes`.

## Variables

Here we create some variables we'll use throughout this process. We create them here and now for convenience.

```{r variables}


```

```{r split-tagging}

# head(Tagging[["Sturgeon"]])

catch <- Split(data = Tagging[["Sturgeon"]], splitVars = RelYear)

# for desired data type
catch$RelYear <- as.numeric(catch[["RelYear"]])

```


## Tag (Mark)

```{r tag-count}

# to get count of White Sturgeon disc-tagged by desired category

# NLeg = legal sized at time of tagging
# NSlt = within current slot (102-152 cm FL); may be variations due to
#        conversions & rounding
# NRng = between 100 & 200 cm TL (range we typically disc tag)

catch$TagCount <- t(vapply(catch[["Data"]], FUN = function(d) {
  b <- d[["Species"]] %in% "White" & d[["StuType"]] %in% "Tag"
  tl <- Filter(f = Negate(is.na), x = d[b, "TL"])
  nall <- sum(b)
  nleg <- sum(d[b, "LenCat"] %in% "leg")
  nslt <- sum(d[b, "InSlot"], na.rm = TRUE)
  nrng <- sum(tl >= 100 & tl <=  200)
  c(NAll = nall, NLeg = nleg, NSlt = nslt, NRng = nrng)
}, numeric(4L)))
# end vapply

```

## Catch

```{r catch-cutoff-mat}

catch$CutoffMat <- CatchCutoffMatrix(
  relYear = unique(Tagging[["Sturgeon"]][["RelYear"]]),
  minLen = c(rep(102, length = 9), 107, 112, rep(117, length = 20)),
  maxLen = c(rep(NA, length = 9), rep(183, length = 10), rep(168, length = 12)),
  vb = GetVBGM(Linf = 261.2, K = 0.04027, t0 = -3.638)
)

```

```{r length-count}

catch$LenCount <- lapply(catch[["Data"]], FUN = function(d) {
  
  b <- d[["Species"]] %in% "White"
  y <- unique(d[["RelYear"]])
  
  clc <- CatchLengthCutoff(catch[["CutoffMat"]], year = y)
  f <- clc[["CountCatch"]]
  f(l = d[b, "TL"])
  
})

# if what = rbind then perform column sums, if what = cbind perform row sums;
# otherwise can use count_lens as a list to process catch per release year

# do.call(what = cbind, args = catch[["LenCount"]])

```

```{r catch-c}

# to get CC (i.e., `C` in MC / R) for each release (tagging) year; not sure if
# removal of NA is needed at this point (05-Aug-2020)

catch$CC <- lapply(catch[["RelYear"]], FUN = function(y) {
  
  o <- vapply(catch[["LenCount"]], FUN = function(x) {
    x[names(x) %in% as.character(y)]
  }, FUN.VALUE = numeric(1L))
  # end vapply
  
  # Filter(f = Negate(is.na), x = o)
  o
  
})
# end outer lapply

```

## Recaptures

```{r recaps}

# CountRecap
# GetStuRecap
# class(Tagging$TagRecaptures$Data$B4010)
# CountRecap(recaps[["CrossTab"]], n = 10)
# dim(recaps$GetMatrixData())

# list to define recapture type (rtype); minDAL is minimum days at large, where
# 30 days ensures adequate mixing with non-tagged populations; can add other
# items to the list as desired (05-Aug-2020)

# defaults: species = 'w', lencat = NULL, minDAL = -1L
rtype <- list(
  All = c(lencat = NULL, minDAL = -1L),
  Leg = c(lencat = 'l', minDAL = 30)
)

# to get summary of recaps from raw data
recap_summary <- lapply(rtype, FUN = function(...) {
  ApplyStuRecap(Tagging[["TagRecaptures"]][["Data"]], ...)
})

# annual (release year) recap count by recapture year
recaps <- lapply(recap_summary, FUN = function(r) {
  
  # to apply table() which gets count by recapture year
  mat <- r$GetMatrixData()
  
  lapply(catch[["RelYear"]], FUN = function(y) {
    b <- mat[, "RelY"] %in% y
    table(mat[b, "RecY"])
  })
  # end inner lapply
})
# end out lapply

```

## Petersen Estimates

```{r petersen-ests}

# to calculate Petersen Estimates (abundance [N]) based on catch & recapture
# collected with each tagging season & number of tags release per year; because
# C is established based on what was legal at time of tagging, these PEs are for
# legal-at-time-of-tagging sized fish (06-Aug-2020)
catch$PetEst <- Map(f = function(m, cc, rr) {
  
  # cc <- Filter(f = Negate(is.na), x = cc)
  
  # to select recap for appropriate year related to catch
  b1 <- names(cc) %in% names(rr)
  
  # to remove same year (within season) value where C is NA (i.e., no annual
  # estimate made for same-year sampling)
  b2 <- !is.na(cc)
  
  # sets up vector of 0s to hold appropriate recapture value
  r <- vector(mode = "integer", length = length(cc))
  r[b1] <- rr
  
  # to set up alpha
  a <- 0.05
  
  # for the three variations on Petersen Estimate (pe)
  # (1) peA: N calculated for each year following sampling year
  # (2) peC: N calculated cumulatively for C & R
  # (3) peT: N calculated based on total (to date) C & R
  list(
    peA = PetersenEst(m, CC = cc[b2], RR = r[b2], alpha = a),
    peC = PetersenEst(m, CC = cumsum(cc[b2]), RR = cumsum(r[b2]), alpha = a),
    peT = PetersenEst(m, CC = sum(cc[b2]), RR = sum(r[b2]), alpha = a)
  )
  
}, m = catch[["TagCount"]][, "NLeg"], cc = catch[["CC"]], rr = recaps[["Leg"]])
# end Map (PetersenEst)

```

```{r display-pe}

catch[["PetEst"]]

```

---
CDFW, SportFish Unit  
`r Sys.time()`
